---
title: "Lightning-Sparked Wildfires in the Apalachicola National Forest (ANF)"
output: html_document
editor_options: 
  chunk_output_type: console
---

1. Writing is thinking, 
2. 1 article, 1 question, 1 finding, 
3. it's not a review of the literature, it's an argument about why we need your paper

What is the relationship between dryness of the soil and lightning-sparked wildfire activity in the ANF on the seasonal time scale? Quantifying the statistical relationship between soil dryness and lightning-sparked wildfire activity in the ANF.

What climate factors are associated with soil dryness during the dry season?

The increasing risk of fire in the ANF. Is the risk of fire in the ANF increasing? Why we need this paper: Uses an index that assesses the risk of fire rather than just rainfall deficit. Looks at changes during the fire season.

KDBI assesses the risk of fire. KDBI is computed from daily rainfall and high temperature. These data are available at TLH NWS. Establish that TLH represents well the ANF by showing a correlation map between KDBI computed at TLH and PRISM data across the ANF. The PRISM dataset should not be used to calculate century-long climate trends due to non-climatic variations from to station equipment and location changes, openings and closings, varying observation times, and the use of relatively short-term networks.

https://www.predictiveservices.nifc.gov/outlooks/monthly_seasonal_outlook.pdf
https://iopscience.iop.org/article/10.1088/1748-9326/ab83a7
https://www.sciencedirect.com/science/article/abs/pii/S0048969719318315

Brown, Emily K., Wang, Jiali, and Feng, Yan. US wildfire potential: a historical view and future projection using high-resolution climate data. United States: N. p., 2021. Web. doi:10.1088/1748-9326/aba868.

Abatzoglou and Williams (2016) found that the correlation coefficient between KBDI and burned area over the forests of southwestern United States is 0.6–0.8.

Title: Seasonal frequency and severity of lightning-sparked wildfires and their relationship to soil dryness and climate variability in the Apalachicola National Forest

The ANF contains some of Florida's largest intact natural areas where lightning-sparked fires are a seasonally common occurrence and wildfire frequency and size distributions are not substantially influenced by confounding effects of urbanization, agricultural land or roadways. While this publicly managed natural area deploys various proactive and reactive fire suppression resources, the incidence of large lightning-sparked fires is still largely a climatic and fuels-driven phenomenon. The landscape is among the largest remaining semi-wild areas of fire-maintained long-leaf pine savanna in the Southeast United States. 

R version 4.1.0 (2021-05-18) -- "Camp Pontanezen"

Fire is controlled by two things: ignition and fuel. Climate change impacts both by increasing flammability (the likelihood something will burn) and the availability of dead, dry stuff on the landscape to burn. People and lightning are ignitors. 

Wildfires require a spark and fuel. In the United States, half of wildfires are initiated by lighting. The other half are human-caused — frequently started by power lines, cigarettes, cars, camp fires or arson. In recent years, there’s been an abundance of very dry fuel. Drought and high heat can kill trees and dry out dead grass, pine needles, and any other material on the bottom of the forest floor that act as kindling when a fire sweeps through a forest. Wildfire experts see the signature of climate change in the dryness, high heat and longer fire season that have made these fires more extreme. “We wouldn’t be seeing this giant ramp up in fire activity as fast as it is happening without climate change,” said Park Williams, a climate scientist at UCLA. 

Like air pollution, wildfire smoke — and particularly the concentration of PM 2.5, or particles smaller than 2.5 microns — can affect the respiratory and cardiovascular systems, said Colleen Reid, an environmental epidemiologist and health geographer at the University of Colorado Boulder.

## Get the required packages
```{r}
library(sf)
library(tidyverse)
library(tidycensus)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(tmap)
library(scales)
library(USAboundaries)
```

## Get a boundary file for the ANF

https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip
```{r}
if(!"S_USA.NFSLandUnit" %in% list.files()){
  download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip",
                "S_USA.NFSLandUnit.zip")
unzip("S_USA.NFSLandUnit.zip",
      files = "S_USA.NFSLandUnit")
}

ANF_Boundary.sf <- st_read(dsn = "S_USA.NFSLandUnit") %>%
  filter(NFSLANDU_2 == "Apalachicola National Forest")

st_area(ANF_Boundary.sf) # 2,564 sq. km (634,000 acre)
```

Make a map showing the ANF boundary and the airport location.
```{r}
Airport.sf <- data.frame(Name = c("Regional Airport"), 
                         Latitude = c(30.39306),
                         Longitude = c(-84.35333)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"),
           crs = 4326) %>%
  st_transform(crs = st_crs(ANF_Boundary.sf))

tmap_mode("view")
tm_shape(ANF_Boundary.sf) +
  tm_borders() +
tm_shape(Airport.sf) +
  tm_dots(size = .03, 
          col = "darkgreen")
```

## Get the wildfire data

Severity of fires: Area burned 
Frequency of fires: Number of fires

```{r}
#unzip("RDS-2013-0009.5_GPKG.zip")
#st_layers(dsn = "Data/FPA_FOD_20210617.gpkg")

Fires.sf <- st_read(dsn = "Data/FPA_FOD_20210617.gpkg",
                    layer = "Fires") %>%
  filter(STATE == "FL") %>%
  st_transform(crs = st_crs(ANF_Boundary.sf)) %>%
  st_intersection(ANF_Boundary.sf)
```

These data were collected using funding from the U.S. Government and can be used without additional permissions or fees. If you use these data in a publication, presentation, or other research product please use the following citation:

Short, Karen C. 2021. Spatial wildfire occurrence data for the United States, 1992-2018 [FPA_FOD_20170508]. 5th Edition. Fort Collins, CO: Forest Service Research Data Archive.
https://www.fs.usda.gov/rds/archive/Catalog/RDS-2013-0009.5

From: https://www.fs.usda.gov/rds/archive/products/RDS-2013-0009.5/_metadata_RDS-2013-0009.5.html
`FIRE_SIZE` = Estimate of acres within the final perimeter of the fire.
`FIRE_SIZE_CLASS` = Code for fire size based on the number of acres within the final fire perimeter expenditures (A=greater than 0 but less than or equal to 0.25 acres, B=0.26-9.9 acres, C=10.0-99.9 acres, D=100-299 acres, E=300 to 999 acres, F=1000 to 4999 acres, and G=5000+ acres).

Abstract: This data publication contains a spatial database of wildfires that occurred in the United States from 1992 to 2018. It is the fourth update of a publication originally generated to support the national Fire Program Analysis (FPA) system. The wildfire records were acquired from the reporting systems of federal, state, and local fire organizations. The following core data elements were required for records to be included in this data publication: discovery date, final fire size, and a point location at least as precise as a Public Land Survey System (PLSS) section (1-square mile grid). The data were transformed to conform, when possible, to the data standards of the National Wildfire Coordinating Group (NWCG), including an updated wildfire-cause standard (approved August 2020). Basic error-checking was performed and redundant records were identified and removed, to the degree possible. In addition to incorporating data for 2016-2018, some previously unavailable nonfederal wildfire records for the period 1999-2015 were acquired either directly from the state fire services (NH, NJ) or indirectly from an updated National Association of State Foresters database (AR, AZ, CA, CO, FL, HI, ID, IL, OK, SD) and added. The resulting product, referred to as the Fire Program Analysis fire-occurrence database (FPA FOD), includes 2.17 million geo-referenced wildfire records, representing a total of 165 million acres burned during the 27-year period. Identifiers necessary to link the point-based, final-fire-reporting information to published large-fire-perimeter and operational-situation-reporting datasets are included.

See also NASA burned-area data https://firms.modaps.eosdis.nasa.gov/ Not used.
```{r, eval=FALSE}
Fires2.sf <- st_read(dsn = "NIFC_ANF_firesshapefile",
                     layer = "NIFCboundedfires")

table(year(Fires2.sf$FireDiscov))
```

Make a bar chart of frequency by cause.
```{r}
table(Fires.sf$NWCG_GENERAL_CAUSE)

df <- Fires.sf %>%
  st_drop_geometry() %>%
  group_by(NWCG_GENERAL_CAUSE) %>%
  summarize(nF = n(),
            perF = nF/nrow(Fires.sf))

ggplot(df,
       mapping = aes(y = reorder(NWCG_GENERAL_CAUSE, perF),
                     x = perF,
                     fill = perF)) +
  geom_col() +
  scale_fill_distiller(palette = "Oranges",
                       direction = 1,
                       guide = "none") +
  scale_x_continuous(labels = percent) +
  ylab("") + xlab("") +
  labs(title = "Lightning is the predominant spark for wildfires in the Apalachicola National Forest",
       subtitle = "Based on data from 1992-2018",
       caption = "Data source: Short, Karen (2021)") +
  theme_minimal()
```

About 40% of all wildfires in the ANF are of natural origin. The next leading cause is arson accounting for just over 16% of all wildfires in the forest.

Create an interactive map showing the locations of all fires over the period of record.
```{r}
tm_shape(ANF_Boundary.sf) +
  tm_borders() +
tm_shape(Fires.sf) +
  tm_bubbles(size = "FIRE_SIZE", col = "orange") +
tm_shape(Airport.sf) +
  tm_dots(size = .1, col = "black")
```

Create a new data frame containing only the natural-caused wildfires.
```{r}
NaturalFires.sf <- Fires.sf %>%
  filter(NWCG_GENERAL_CAUSE == "Natural") %>%
  mutate(Year = year(DISCOVERY_DATE),
         Month = month(DISCOVERY_DATE),
         Day = day(DISCOVERY_DATE),
         YDay = yday(DISCOVERY_DATE),
         YDayF = factor(YDay, levels = as.character(1:366)),
         SizeF = factor(FIRE_SIZE_CLASS, ordered = TRUE)) %>%
  dplyr::select(Year, Month, Day, YDay, YDayF, Size = FIRE_SIZE, SizeF)
```

Draw a static map showing the location of the natural-caused wildfires by size of area burned.
```{r}
FL_Counties.sf <- us_counties(states = "FL",
                              resolution = "high") %>%
  st_transform(crs = st_crs(ANF_Boundary.sf)) %>%
  st_crop(st_bbox(st_buffer(ANF_Boundary.sf, dist = .25)) )

ggplot() +
  geom_sf(data = FL_Counties.sf, fill = "transparent", col = "gray80") +
  geom_sf(data = ANF_Boundary.sf, fill = "transparent") +
  geom_sf(data = NaturalFires.sf,
          mapping = aes(col = SizeF)) +
  scale_color_brewer(palette = "Oranges",
                     direction = 1,
                     name = "Size Class") +
  theme_bw() +
  labs(title = "Location of natural-caused0 wildfires in the Apalachicola National Forest (1992-2018)",
       subtitle = "Darker color points indicates the fire resulted in a larger burn area",
       caption = "Data source: Short, Karen (2021)")
```

Are the locations of natural-caused wildfires during May, June, and July within the ANF clustered?

Create a {spatstat} object containing the locations of lightning-sparked fires in the ANF during May-July. Get a planar projection suggested from `suggest_crs()`. Here projected 2779 NAD83(HARN) / Florida North
```{r}
library(spatstat)
library(maptools)
#remotes::install_github("walkerke/crsuggest")
library(crsuggest)

suggest_crs(ANF_Boundary.sf)

W <- ANF_Boundary.sf %>% 
  st_transform(crs = 2779) %>%
  as_Spatial() %>% 
  as.owin()

NF.ppp <- NaturalFires.sf %>% 
  filter(Month %in% c(5, 6, 7)) %>%
  st_geometry() %>%
  st_transform(crs = 2779) %>%
  as_Spatial() %>% 
  as.ppp()

NF.ppp <- NF.ppp[W] %>%
  rescale(s = 1000, 
          unitname = "km")
summary(NF.ppp)
```

Area is 2,564 square km. There are 356 natural wildfires for an average intensity of 13.9 fires per 10 square kilometer during May-July over the 27-year period.
```{r}
NF.ppp %>%
  density() %>%
  plot()

NF.ppp %>%
  dirichlet() %>%
  plot()

NF.ppp %>%
  quadratcount(nx = 20, ny = 15) %>%
  plot()

NF.ppp %>%
  quadrat.test(nx = 20, ny = 15,
               method = "M",
               nsim = 9999,
               alternative = "clustered")
```

There is some weak evidence that the locations are clustered. But the evidence is not significant.

Compute Ripley K using the `Kstat()` function
```{r}
df <- NF.ppp %>%
  envelope(fun = Kest,
           nsim = 100,
           global = TRUE,
           nrank = 1) %>%
  as.data.frame()

df <- df %>%
  mutate(theo = theo * intensity(NF.ppp),
         obs = obs * intensity(NF.ppp),
         lo = lo * intensity(NF.ppp),
         hi = hi * intensity(NF.ppp))

ggplot(data = df,
       mapping = aes(x = r, y = obs)) +
  geom_ribbon(aes(ymin = lo, 
                  ymax = hi), fill = "gray80") +
  geom_line(col = "orange") +
  geom_line(aes(y = theo), color = "white") +
  xlab("Distance (km)") + ylab("K(r)") +
  labs(title = "Locations of natural wildfires in the ANF are inconsistent with a random process",
       subtitle = "Average observed (orange) versus modeled (white) number of additional wildfires within a distance r of a random wildfire.\nThe gray ribbon is the 99% uncertainy band based on 100 simulations from the spatial random process model.") +
  theme_minimal()
```

At larger scales there are features that make natural fires somewhat more or less likely (e.g., Bradwell Bay Wilderness--swamp) 

Monthly occurrence (relative frequency) of natural wildfires in the ANF.
```{r}
df <- NaturalFires.sf %>%
  st_drop_geometry() %>%
  mutate(MonthF = factor(month.name[Month], 
                         levels = rev(month.name), 
                         ordered = TRUE)) %>%
  group_by(MonthF, .drop = FALSE) %>%
  summarize(nF = n(),
            perF = nF/nrow(NaturalFires.sf))

ggplot(data = df,
       mapping = aes(y = MonthF, 
                     x = perF,
                     fill = perF)) +
  geom_col() +
  scale_fill_distiller(palette = "Oranges",
                       direction = 1,
                       guide = "none") +
  scale_x_continuous(labels = percent) +
  xlab("") + ylab("") +
  labs(title = "Over 80% of natural wildfires in the Apalachicola National Forest occur during May-July",
       subtitle = "Percentage of all natural wildfires by month",
       caption = "Period of record: 1992-2018, Data source: Short, Karen (2021)") +
  theme_minimal() 
```

Seasonal number of natural fires and seasonal total acres burned.
```{r}
df <- NaturalFires.sf %>%
  st_drop_geometry() %>%
  filter(Month %in% c(5, 6, 7)) %>%
  mutate(YearF = factor(Year, levels = as.character(1992:2018), ordered = TRUE)) %>%
  group_by(YearF, .drop = FALSE) %>%
  summarize(nFires = n(),
            totalAcresBurned = sum(Size)) %>%
  mutate(Year = 1992:2018)

ggplot(data = df,
       mapping = aes(x = Year, y = nFires, color = totalAcresBurned)) +
  geom_line(col = "gray80") +
  geom_point(mapping = aes(size = totalAcresBurned)) +
  scale_size_continuous(guide = "none") +
  scale_color_distiller(palette = "Oranges",
                        direction = 1,
                        guide = "none") +
  scale_x_continuous(breaks = seq(1992, 2018, by = 3)) +
  theme_minimal() +
  labs(x = "", y = "",
       title = "Large inter-seasonal variability in the number of natural wildfires in the ANF",
       subtitle = "Larger and darker points indicate larger burn area")

ggplot(data = df,
       mapping = aes(x = Year, y = totalAcresBurned, color = nFires)) +
  geom_line(col = "gray80") +
  geom_point(mapping = aes(size = nFires)) +
  scale_size_continuous(guide = "none") +
  scale_color_distiller(palette = "Oranges",
                        direction = 1,
                        guide = "none") +
  scale_x_continuous(breaks = seq(1992, 2018, by = 3)) +
  theme_minimal() +
  labs(x = "", y = "",
       title = "Large inter-seasonal variability in acres burned from natural wildfires in the ANF",
       subtitle = "Larger and darker color points indicate more fires")
```

Seasonal temporal autocorrelation in frequency and total area burned.
```{r}
plot(acf(df$nFires))
plot(acf(df$totalAcresBurned))
cor.test(df$nFires, df$totalAcresBurned)
cor.test(df$nFires[-1], df$totalAcresBurned[-24])
```

No significant inter-seasonal autocorrelation. Significant correlation between number of fires and burned area as expected.

Possible project: Make fire size an ordered factor and create a model that predicts the probability by size category similar to EF rating.
```{r}
NaturalFires.sf %>%
  group_by(SizeF) %>%
  summarize(nFires = n(),
            Range1 = range(Size)[1],
            Range2 = range(Size)[2])
```

https://www.nifc.gov/nicc/predictive/outlooks/outlooks.htm
https://www.cpc.ncep.noaa.gov/products/predictions/90day/

## Get lightning data

Daily county-level counts 1986-2013. Data location: https://www1.ncdc.noaa.gov/pub/data/swdi/reports/county/byFips/

First get the Florida FIPS codes for Liberty, Wakulla, and Franklin counties, then get the data.
```{r}
FLfips <- fips_codes %>%
  filter(state == "FL") %>%
  filter(county %in% c("Liberty County", "Wakulla County")) %>%
  pull(county_code)

fn <- paste0("https://www1.ncdc.noaa.gov/pub/data/swdi/reports/county/byFips/swdireport-12", FLfips, "-BETA.csv")

df <- data.frame()
for(i in 1:length(fn)){
  X <- read.csv(fn[i], na.strings = "NULL", header = TRUE, stringsAsFactors = FALSE)
  df <- rbind(df, X)
}

df <- df %>%
  mutate(SEQDAY = as.Date(SEQDAY),
         DAY = day(SEQDAY),
         MONTH = month(SEQDAY),
         YEAR = year(SEQDAY))
```

Seasonal cycle.
```{r}
sc <- df %>%
  mutate(MonthF = factor(month.name[MONTH], 
                         levels = rev(month.name), 
                         ordered = TRUE)) %>%
  group_by(MonthF, .drop = FALSE) %>%
  summarize(AvgDailyCount = mean(FCOUNT_NLDN, na.rm = TRUE))

ggplot(data = sc,
       mapping = aes(y = MonthF, 
                     x = AvgDailyCount,
                     fill = AvgDailyCount)) +
  geom_col() +
  scale_fill_distiller(palette = "Blues",
                       direction = 1,
                       guide = "none") +
  labs(x = "", y = "",
       title = "Lightning in the Apalachicola National Forest peaks from June-August",
       subtitle = "Daily average number of cloud-to-ground lightning strikes by month",
       caption = "Period of record: 1986-2013, Data source: NCEI") +
  theme_minimal() 
```

Cloud to ground lightning is most common in the ANF between June through August. The wildfire season beats to the rhythms of drying and lightning.

## Get TLH airport daily weather data and compute the Keetch & Byram drought index

Original paper outlining the rationale and how to create it: https://www.srs.fs.usda.gov/pubs/rp/rp_se038.pdf

The Keetch-Byram Drought Index assesses the risk of fire by representing the net effect of evapotranspiration and precipitation in producing cumulative moisture deficiency in deep duff and upper soil layers. The index ranges from zero, the point of no moisture deficiency, to 800, the maximum drought that is possible. Units are 100's of inches. KBDI relates current and recent weather conditions to potential or expected fire behavior. It was originally developed for the Southeast United States and is based primarily on recent rainfall patterns. It is one of the only drought indexes specifically developed to equate the effects of drought with potential fire behavior. KBDI is used for fire studies as opposed to hydrological drought indices: https://piahs.copernicus.org/articles/383/291/2020/ 

For different soil types, the depth of soil required to hold 8 inches of moisture varies (loam = 30", clay = 25" and sand = 80"). A prolonged drought (high KBDI) influences fire intensity largely because more fuel is available for combustion (i.e. fuels have a lower moisture content). In addition, the drying of organic material in the soil can make it harder to suppress fires.

High values of the KBDI indicate conditions favorable for the occurrence and spread of wildfires, but drought is not by itself a prerequisite for wildfires. Other weather factors, such as wind, temperature, relative humidity and atmospheric stability, play a major role in determining the actual fire danger.

NWS Tallahassee daily weather data:
https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USW00093805/detail
https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/readme.txt
Column explanations: https://docs.google.com/document/d/1q2WEpXndpMx9lUq-0ON63GojkaOzixrGyYEhI_xkPtw/edit?usp=sharing
AWND = Average daily wind speed (tenths of meters per second) 

Import the summary of the day data and add columns to the data frame. Assign zero rainfall on days with missing values.
```{r}
TLH.df <- read.csv(file = 'Data/TLH_Daily1940-2020.csv',
                   stringsAsFactors = FALSE,
                   header = TRUE) %>%
  mutate(Date = as.Date(DATE)) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24mm = Rainfall24 * 25.4) %>%
  filter(Year <= 2020)

sum(is.na(TLH.df$PRCP))
sum(is.na(TLH.df$PRCP)) / nrow(TLH.df) * 100  # only 5 missing values [~ .017% of all days]
sum(is.na(TLH.df$TMAX))
sum(is.na(TLH.df$TMAX)) / nrow(TLH.df) * 100  # only 1 missing value [~ .0034% of all days]

# Missing temperature and rainfall filled in from NCDC daily summaries
# https://gis.ncdc.noaa.gov/maps/ncei/summaries/daily

TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 95
TLH.df$Rainfall24[is.na(TLH.df$PRCP)] <- c(.1, .3, 0, .1, 0)
TLH.df$Rainfall24mm[is.na(TLH.df$PRCP)] <- c(3, 8.9, 0, 2.5, 0)

( AnnualRainfall <- TLH.df %>%
  filter(Year >= 1943) %>%  # complete data started in March 1943
  group_by(Year) %>%
  summarize(totalRainfall = sum(Rainfall24)) %>%
  pull(totalRainfall) %>%
  mean() )
```

Plot monthly values of average daily high temperature and average month rainfall.
```{r}
df <- TLH.df %>%
  filter(Year >= 1943 & Year <= 2019) %>%
  group_by(month) %>%
  summarize(totalRain = sum(Rainfall24mm),
            avgMonthlyTotalRain = totalRain/76,
            avgDailyHighTemp = mean(MaxTemp))

p1 <- ggplot(data = df,
             mapping = aes(x = month, y = avgMonthlyTotalRain, fill = avgMonthlyTotalRain)) +
        geom_col() +
        scale_fill_distiller(palette = "Greens",
                         direction = 1, 
                         guide = 'none') +
        theme_dark() +
        labs(x = "", y = "(mm)",
             title = "Average monthly rainfall amount in the Apalachicola National Forest",
             subtitle = "1943-2020")


p2 <- ggplot(data = df,
             mapping = aes(x = month, y = avgDailyHighTemp, color = avgDailyHighTemp)) +
        geom_point(size = 4) +
        scale_color_distiller(palette = "Oranges",
                              direction = 1, 
                              guide = 'none') +
        scale_y_continuous(limits = c(60, 100)) +
        theme_dark() +
        labs(x = "", y = "(°F)",
             title = "Average daily high temperature in the Apalachicola National Forest",
             subtitle = "1943-2020")

library(patchwork)

p1 / p2

```

Compute daily values of the KBDI. Original paper outlining the rationale and how to create it: https://www.srs.fs.usda.gov/pubs/rp/rp_se038.pdf with a minor correction identified in Alexander1992.pdf (paper found in Dropbox/Literature).

Step one: Compute net rainfall on each day. Units are inches.
```{r}
Rainfall24 <- TLH.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0

CumR <- 0
NetR <- numeric()

for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH.df$NetR <- NetR
```

Step two: Compute daily drought index. Units of temperature are degrees F.
```{r}
Q <- 500  # the influence of this start value diminishes to zero after a few months
R <- AnnualRainfall # average annual rainfall for TLH in inches

MaxTemp <- TLH.df$MaxTemp

Ql <- numeric()
DeltaQl <- numeric()
for(i in 1:length(Rainfall24)){
  DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
  Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
  Q <- ifelse(Q < 0, 0, Q) 
  Ql <- c(Ql, Q)
  DeltaQl <- c(DeltaQl, DeltaQ)
}

TLH.df$Ql <- Ql
TLH.df$Qlm <- Ql * .254  # tenth of an inch to mm
TLH.df$DeltaQl <- DeltaQl
TLH.df$DroughtIndex <- floor(Ql/100)

TLH.df <- TLH.df %>%
 dplyr::filter(Year >= 1943 & Year <= 2020) # Only full years after the daily Q values have converged
range(TLH.df$Ql)
```

Note: There is a package to compute the standardized precip-evapotranspiration index {SPEI}. Only good for monthly data. Same for the Palmer Drought Severity Index (PDSI).

Monthly average soil moisture deficit.
```{r}
TLH.df %>%
  mutate(MonthF = factor(month.name[month], 
                         levels = rev(month.name), 
                         ordered = TRUE)) %>%
  group_by(MonthF, .drop = FALSE) %>%
  summarize(AvgSoilMoistureDeficit = mean(Qlm)) %>%
ggplot(mapping = aes(y = MonthF, 
                     x = AvgSoilMoistureDeficit,
                     fill = AvgSoilMoistureDeficit)) +
  geom_col() +
#  scale_fill_distiller(palette = "BrBG",
#                       direction = -1,
#                       guide = 'none') +
  scale_fill_gradientn(colors = terrain.colors(5),
                       guide = 'none') +
#  scale_fill_gradient2(low = "#FC8B93", mid = "#965784", high = "#D9DE6E", guide = 'none') +
  labs(x = "", y = "",
       title = "May through November is the dry season in the Apalachicola National Forest",
       subtitle = "Average soil moisture deficit (mm)",
       caption = "Period of record: 1943-2020, Data source: NWSFO Tallahassee") +
  theme_dark() 
```

Make some trend plots. The second plot is statistically more realistic because it removes the temporal autocorrelations.
```{r}
ggplot(TLH.df, aes(x = Date, y = Qlm, color = Qlm)) +
  geom_line() +
  scale_color_gradientn(colors = terrain.colors(5), guide = 'none') +
  ylab("") + xlab("") +
  geom_smooth(method = lm, se = FALSE, color = "white") +
  facet_wrap(~ month, ncol = 12) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Amount of water needed to bring soil moisture to full capacity in the Apalachicola National Forest",
       subtitle = "Daily soil moisture deficit (mm)", 
       caption = "Period of record: 1943-2020, Data source: NWSFO Tallahassee") 

TLH.df %>%
  group_by(month, Year) %>%
  summarise(Avg = mean(Qlm)) %>%
  ggplot(aes(x = Year, y = Avg, color = Avg)) +
  geom_smooth(method = lm, se = FALSE, color = "gray70") +
  geom_point() +  
  scale_color_gradientn(colors = terrain.colors(5), guide = "none") +
  scale_y_continuous(limits = c(0, 201)) +
  scale_x_continuous(limits = c(1940, 2020), breaks = c(1950, 1980, 2010)) +
  ylab("") + xlab("") +
  facet_wrap(~ month, ncol = 12) +
  theme_dark() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Risk of wildfires is increasing in the Apalachicola National Forest.",
       subtitle = "Monthly average soil moisture deficit (mm) by year with trend line (gray)", 
       caption = "Period of record: 1943-2020, Data source: NWSFO Tallahassee") 
```

Side-by-side daily distribution plots (early vs late) by month. 1946:1980, 1981:2019
```{r, eval=FALSE}
TLH.df %>%
  mutate(Epoch = Year > 1980) %>%
  ggplot(aes(x = Qlm, color = Epoch)) +
  geom_freqpoly(size = 1, bins = 8) +
  scale_color_manual(labels = c("1943-1980", "1981-2020"), values = c("black", "sandybrown")) +
  scale_x_continuous(limits = c(0, 250), breaks = c(0, 100, 200)) +
  ylab("Number of days") + xlab("Daily soil moisture deficit (mm)") +
  facet_wrap(~ month, ncol = 12) +
  theme_minimal()
```

Repeat using data from Apalachicola (see `compare.rmd`). https://www.ncdc.noaa.gov/  > I want to search for data at a particular location > DATA TOOLS > Final a Station

Linear trend model. Statistical significance of the trend by month.
```{r}
summary(lm(Qlm ~ Date, data = TLH.df)) # this model assumes daily Qlm are independent (they are certainly not)

library(broom)

( Trends.df <- TLH.df %>% 
  group_by(Year, month) %>%
  summarize(AvgQlm = mean(Qlm)) %>%
  group_by(month) %>%
  do(tidy(lm(AvgQlm ~ Year, data = .))) %>%
  dplyr::filter(term == "Year") )

Avg <- TLH.df %>%
  group_by(month) %>%
  summarize(AvgQlm = mean(Qlm)) %>%
  pull(AvgQlm)

cor(Trends.df$estimate, Avg) # positive, indicating upward trends are occurring in months with higher drought index values

Avg <- TLH.df %>%
  group_by(month) %>%
  summarize(nYear = n_distinct(Year), 
            TotalPRCP = sum(PRCP, na.rm = TRUE),
            AvgMonthlyPRCP = TotalPRCP/nYear) %>%
  pull(AvgMonthlyPRCP)

cor(Trends.df$estimate, Avg)  # negative as expected since upward trends are occurring in months with less rainfall
```

Soil moisture deficits are increasing at .4 mm/year on average during April. Which amounts to 4 mm/decade.

What is the correlation between PRCP, TMAX and Qlm grouped by month?
```{r}
TLH.df %>%
  group_by(month) %>%
  summarize(rPRCP = cor(Qlm, PRCP, use = "complete"),
            rTMAX = cor(Qlm, TMAX, use = "complete"),
            rTMIN = cor(Qlm, TMIN, use = "complete"),
            rTAVG = cor(Qlm, TAVG, use = "complete"))
```

Seasonal (fire season) temperature vs KDBI scatter plot. Create a character variable called `FireSeason` (May, June, July). Then group by that variable computing the average temp and total precipitation (or avg KDBI).
```{r}
df <- TLH.df %>%
  mutate(FireSeason = month %in% c("May", "Jun", "Jul")) %>%
  filter(FireSeason) %>%
  group_by(Year) %>%
  summarize(AvgHiTemp = mean(MaxTemp),
            AvgLoTemp = mean(MinTemp),
            TotalPrcp = sum(Rainfall24mm),
            TotalNetR = sum(NetR),
            AvgQlm = mean(Qlm))

ggplot(df, 
       mapping = aes(y = TotalNetR, x = AvgHiTemp, label = Year, color = factor(Year))) +
  geom_point() +
  geom_label_repel(size = 2.5) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_color_grey(start = .8, end = .2, guide = "none") +
#  geom_smooth(method = lm, se = FALSE, col = "black") +
  ylab("Total Precipitation (mm)") + xlab("Average Daily High Temperature (°F)") +
  ggtitle(label = "Fire seasons in the Apalachicola National Forest (May-Jul) that are drier tend to be hotter",
          subtitle = "Dry days are sunnier and thus they tend to get hotter. Later years have darker labels.") +
  theme_minimal() 
```

Get SOI data and join with TLH.df.
```{r}
SOI.df <- read.csv(file = "https://www.ncdc.noaa.gov/teleconnections/enso/indicators/soi/data.csv",
                   skip = 1, header = TRUE) %>%
  mutate(Date = parse_date_time(as.character(Date), "ym"),
         Year = year(Date),
         Month = month(Date),
         month = month(Date, label = TRUE, abbr = TRUE),
         SOI = Value) %>%
  dplyr::select(Year, Month, month, SOI)

Joined.df <- TLH.df %>% 
  dplyr::filter(Year >= 1951) %>%
  group_by(Year, month) %>%
  summarize(AvgQlm = mean(Qlm)) %>%
  dplyr::select(Year, month, AvgQlm) %>%
  left_join(SOI.df, by = c("Year", "month"))

Joined.df %>%
  group_by(month) %>%
  do(tidy(lm(AvgQlm ~ Year + SOI, data = .))) %>%
  dplyr::filter(term == "Year")

Joined.df %>%
  group_by(month) %>%
  do(tidy(lm(AvgQlm ~ Year + SOI, data = .))) %>%
  dplyr::filter(term == "SOI")
```

SOI is positively correlated with moisture deficit. La Nina conditions (positive SOI) during winter into spring leads to larger moisture deficits (higher Qlm). The upward trend in Qlm remains significant after accounting for SOI.

In terms of frequency by Drought Index category. Proportion of the year in each drought category.
```{r}
TLH.df %>%
  group_by(Year, DroughtIndex) %>%
  summarize(N = n()) %>%
  ggplot(aes(x = Year, y = N)) +
  geom_col(aes(fill = as.factor(DroughtIndex))) +
  scale_fill_manual(values = terrain.colors(8), name = "KDBI") +
  scale_x_continuous(name = "", breaks = seq(1941, 2020, 4)) +
  scale_y_reverse() +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(title = "The proportion of the year in drought (KDBI > 5) in the ANF has increased since 1960",
       subtitle = "KDBI = 7-6: Drought, 4-5: Dry, 1-3: Moist, 0: Saturated",
       caption = "Period of record: 1941-2020, Data source: NWSFO Tallahassee") 
```

Daily
```{r}
ggplot(data = TLH.df, 
       mapping = aes(x = doy, y = Year, fill = Qlm)) +
  geom_tile() +
  scale_y_reverse(breaks = seq(1941, 2020, 4)) +
  scale_x_continuous(breaks = c(15,	46,	75,	106, 136, 167, 197, 228, 259, 289, 320, 350),
                     labels = month.abb, position = "top") +
#  scale_fill_gradient2(name = "", low = "#FC8B93", mid = "#965784", high = "#D9DE6E") +
  scale_fill_gradientn(colors = terrain.colors(8), name = "mm") +
  ylab("") + xlab("") +
  theme_minimal() +
  labs(title = "Amount of water needed to bring soil moisture to full capacity in the Apalachicola National Forest", 
      subtitle = "Daily soil moisture deficit (mm)",
      caption = "Period of record: 1941-2020, Data source: NWSFO Tallahassee") 

ggplot(data = TLH.df, 
       mapping = aes(x = doy, y = Year, fill = factor(DroughtIndex))) +
  geom_tile() +
  scale_y_reverse(breaks = seq(1948, 2019, 4)) +
  scale_x_continuous(breaks = c(15,	46,	75,	106, 136, 167, 197, 228, 259, 289, 320, 350),
                     labels = month.abb, position = "top") +
  scale_fill_ordinal(name = "", direction = 1, alpha = .6) +
  ylab("") + xlab("") +
  theme_minimal() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(title = "Daily KBDI category in the Apalachicola National Forest",
       subtitle = "KDBI = 7-6: Drought, 4-5: Dry, 1-3: Moist, 0: Saturated",
       caption = "Period of record: 1949-2019, Data source: NWSFO Tallahassee")
```

Number of days per year with KBDI above 650.
```{r}
library(MASS)

TLH.df %>%
  group_by(Year) %>%
  summarize(nD = sum(Ql > 650)) %>%
  ggplot(aes(Year, nD)) +
  geom_point() +
  stat_smooth(method = "glm.nb",
              formula = y ~ x, 
              se = FALSE,
              col = "orange") 
```

Combine seasonal fire data with drought data.
```{r}
SeasonalFires <- LightningFires.sf %>%
  st_drop_geometry() %>%
  filter(Month %in% c(5, 6, 7)) %>%
  mutate(YearF = factor(Year, levels = as.character(1992:2015), ordered = TRUE)) %>%
  group_by(YearF, .drop = FALSE) %>%
  summarize(nFires = n(),
            totalAcresBurned = sum(FIRE_SIZE)) %>%
  mutate(Year = 1992:2015)

Drought <- TLH.df %>%
  dplyr::filter(Year >= 1992 & Year <= 2015) %>%
  dplyr::filter(month == "Apr") %>%
  group_by(Year) %>%
  summarize(QlmLastDay = last(Qlm),
            AvgQlm = mean(Qlm))

AprilSOI <- SOI.df %>%
  dplyr::filter(Year >= 1992 & Year <= 2015) %>%
  dplyr::filter(month == "Apr") %>%
  pull(SOI)

df <- data.frame(Year = Drought$Year, 
                 QlmLastDay = Drought$QlmLastDay,
                 AvgQlm = Drought$AvgQlm,
                 nFires = SeasonalFires$nFires, 
                 totalAcresBurned = SeasonalFires$totalAcresBurned,
                 AprilSOI = AprilSOI)

write_csv(df, file = "SeasonalForecastModelData.csv")
```

Correlations and scatter plots.
```{r}
cor(df$AvgQlm, df$totalAcresBurned) # .57
cor(df$QlmLastDay, df$totalAcresBurned) #.52
cor(df$QlmLastDay, df$nFires) # .63
cor(df$AvgQlm, df$nFires) # .58

ggplot(data = df,
       mapping = aes(x = QlmLastDay, y = nFires, label = Year)) +
  geom_point() +
  geom_label_repel() +
  theme_minimal()
```

Correlation by day of month during April.
```{r}
TLH.df %>%
  dplyr::filter(Year >= 1992 & Year <= 2015) %>%
  dplyr::filter(month == "Apr") %>%
  group_by(doy, Year) %>%
  mutate(Qlm = Qlm) %>%
  dplyr::select(Year, doy, Qlm) %>%
  filter(doy != 91 & doy != 121) %>%
  group_by(doy) %>%
  summarize(r = cor.test(Qlm, SeasonalFires$nFires)$estimate,
            rLo = cor.test(Qlm, SeasonalFires$nFires, conf.level = .9)$conf.int[1],
            rHi = cor.test(Qlm, SeasonalFires$nFires, conf.level = .9)$conf.int[2]) %>%
  ggplot(aes(x = doy, y = r)) +
  geom_ribbon(aes(ymin = rLo, ymax = rHi), fill = "gray90") +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(-1, 1)) +
#  scale_x_continuous(breaks = c(1,	32,	60,	91, 121, 152, 182, 213, 244, 274, 305, 335),
#                     labels = month.abb, position = "top") +
  theme_minimal()
```

A slight upward trend in the correlation from early to late April but no significant difference.

Continue with `PredictionModel.Rmd` and `GriddedClimateData.Rmd`.