---
title: "Lightning-Initiated Fires in the Apalachicola National Forest (ANF)"
output: html_document
editor_options: 
  chunk_output_type: console
---

Title: Predicting the frequency and severity of lightning-initiated wildfires in the Apalachicola National Forest

The ANF contains some of Florida's largest intact natural areas where lightning-initiated fires are a seasonally common occurrence and wildfire frequency and size distributions are not substantially influenced by confounding effects of urbanization, agricultural land or roadways. While such publicly managed natural areas deploy various proactive and reactive fire suppression resources, the incidence of large lightning-initiated fires is still largely a climatic and fuels-driven phenomenon. These landscapes are among the largest remaining semi-wild areas of fire-maintained long-leaf pine savanna in the Southeast United States. 

R version 4.0.5 (2021-03-31) -- "Shake and Throw"

## Get the required packages
```{r}
library(sf)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(tmap)
library(scales)
library(USAboundaries)
```

## Get a boundary file for the ANF

https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip
```{r}
if(!"S_USA.NFSLandUnit" %in% list.files()){
  download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip",
                "S_USA.NFSLandUnit.zip")
unzip("S_USA.NFSLandUnit.zip",
      files = "S_USA.NFSLandUnit")
}

ANF_Boundary.sf <- st_read(dsn = "S_USA.NFSLandUnit") %>%
  filter(NFSLANDU_2 == "Apalachicola National Forest")

st_area(ANF_Boundary.sf) # 2,564 sq. km (634,000 acre)
```

Make a map showing the ANF boundary and the airport location.
```{r}
Airport.sf <- data.frame(Name = c("Regional Airport"), 
                         Latitude = c(30.39306),
                         Longitude = c(-84.35333)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"),
           crs = 4326) %>%
  st_transform(crs = st_crs(ANF_Boundary.sf))

tmap_mode("view")
tm_shape(ANF_Boundary.sf) +
  tm_borders() +
tm_shape(Airport.sf) +
  tm_dots(size = .03, 
          col = "darkgreen")
```

## Get the wildfire data

Severity of fires: Area burned 
Frequency of fires: Number of fires

See also `Get_Wildfires.Rmd` on Dropbox. These data were collected using funding from the U.S. Government and can be used without additional permissions or fees. If you use these data in a publication, presentation, or other research product please use the following citation:

Short, Karen C. 2017. Spatial wildfire occurrence data for the United States, 1992-2015 [FPA_FOD_20170508]. 4th Edition. Fort Collins, CO: Forest Service Research Data Archive. https://doi.org/10.2737/RDS-2013-0009.4

From: https://www.fs.usda.gov/rds/archive/products/RDS-2013-0009.4/_metadata_RDS-2013-0009.4.html
`FIRE_SIZE` = Estimate of acres within the final perimeter of the fire.
`FIRE_SIZE_CLASS` = Code for fire size based on the number of acres within the final fire perimeter expenditures (A=greater than 0 but less than or equal to 0.25 acres, B=0.26-9.9 acres, C=10.0-99.9 acres, D=100-299 acres, E=300 to 999 acres, F=1000 to 4999 acres, and G=5000+ acres).

Abstract: This data publication contains a spatial database of wildfires that occurred in the United States from 1992 to 2015. It is the third update of a publication originally generated to support the national Fire Program Analysis (FPA) system. The wildfire records were acquired from the reporting systems of federal, state, and local fire organizations. The following core data elements were required for records to be included in this data publication: discovery date, final fire size, and a point location at least as precise as Public Land Survey System (PLSS) section (1-square mile grid). The data were transformed to conform, when possible, to the data standards of the National Wildfire Coordinating Group (NWCG). Basic error-checking was performed and redundant records were identified and removed, to the degree possible. The resulting product, referred to as the Fire Program Analysis fire-occurrence database (FPA FOD), includes 1.88 million geo-referenced wildfire records, representing a total of 140 million acres burned during the 24-year period.
```{r}
Fires.sf <- st_read(dsn = "Data/FL_Fires",
                    layer = "FL_Fires")  %>%
  st_transform(crs = st_crs(ANF_Boundary.sf)) %>%
  st_intersection(ANF_Boundary.sf)
```

Make a bar chart of frequency by cause.
```{r}
table(Fires.sf$STAT_CAU_1)

df <- Fires.sf %>%
  st_drop_geometry() %>%
  group_by(STAT_CAU_1) %>%
  summarize(nF = n(),
            perF = nF/nrow(Fires.sf))

ggplot(df,
       mapping = aes(y = reorder(STAT_CAU_1, perF),
                     x = perF,
                     fill = perF)) +
  geom_col() +
  scale_fill_distiller(palette = "Oranges",
                       direction = 1,
                       guide = FALSE) +
  scale_x_continuous(labels = percent) +
  ylab("") + xlab("") +
  labs(title = "Lightning is the predominant spark for wildfires in the Apalachicola National Forest",
       subtitle = "Based on data from 1992-2015",
       caption = "Data source: Short, Karen (2017)") +
  theme_minimal()
```

Locations of all fires over the period of record.
```{r}
tm_shape(ANF_Boundary.sf) +
  tm_borders() +
tm_shape(Fires.sf) +
  tm_bubbles(size = "FIRE_SIZE", col = "orange") +
tm_shape(Airport.sf) +
  tm_dots(size = .1, col = "black")
```

Lightning caused fires.
```{r}
LightningFires.sf <- Fires.sf %>%
  filter(STAT_CAU_1 == "Lightning") %>%
  mutate(Year = year(DISCOVERY_),
         Month = month(DISCOVERY_),
         Day = day(DISCOVERY_),
         YDay = yday(DISCOVERY_),
         YDayF = factor(YDay, levels = as.character(1:366)),
         SizeF = factor(FIRE_SIZE_, ordered = TRUE)) %>%
  select(Year, Month, Day, YDay, YDayF, FIRE_SIZE, SizeF)
```

Map of lightning caused fires by size of the area burned.
```{r}
FL_Counties.sf <- us_counties(states = "FL",
                             resolution = "high") %>%
  st_transform(crs = st_crs(ANF_Boundary.sf)) %>%
  st_crop(st_bbox(st_buffer(ANF_Boundary.sf, dist = .25)) )

ggplot() +
  geom_sf(data = FL_Counties.sf, fill = "transparent", col = "gray80") +
  geom_sf(data = ANF_Boundary.sf, fill = "transparent") +
  geom_sf(data = LightningFires.sf,
          mapping = aes(col = SizeF), show.legend = FALSE) +
  scale_color_brewer(palette = "Oranges",
                     direction = 1,
                     guide = FALSE) +
  theme_bw() +
  labs(title = "Location of lightning-caused wildfires in the Apalachicola National Forest (1992-2015)",
       subtitle = "Darker color points indicates the fire resulted in a larger burn area",
       caption = "Data source: Short, Karen (2017)")
```

Project: create a {spatstat} object and compute Ripley K using the `Kstat()` function. Get a planar projection suggested from `suggest_crs()`. Here projected 2779 NAD83(HARN) / Florida North
```{r}
library(spatstat)
library(maptools)
#remotes::install_github("walkerke/crsuggest")
library(crsuggest)

suggest_crs(ANF_Boundary.sf)

W <- ANF_Boundary.sf %>% 
  st_transform(crs = 2779) %>%
  as_Spatial() %>% 
  as.owin()

LF.ppp <- Fires.sf %>% 
  filter(STAT_CAU_1 == "Lightning") %>%
  mutate(Month = month(DISCOVERY_)) %>%
  filter(Month %in% c(5, 6, 7)) %>%
  st_geometry() %>%
  st_transform(crs = 2779) %>%
  as_Spatial() %>% 
  as.ppp()

LF.ppp <- LF.ppp[W] %>%
  rescale(s = 1000, 
          unitname = "km")
summary(LF.ppp)

df <- LF.ppp %>%
  envelope(fun = Kest,
           nsim = 100,
           global = TRUE,
           nrank = 1) %>%
  as.data.frame()

#df <- LF.ppp %>%
#  lohboot(fun = Kest,
#          block = TRUE,
#          nx = 10,
#          global = FALSE) %>%
#  as.data.frame()

df <- df %>%
  mutate(theo = theo * intensity(LF.ppp),
         obs = obs * intensity(LF.ppp),
         lo = lo * intensity(LF.ppp),
         hi = hi * intensity(LF.ppp))

ggplot(data = df,
       mapping = aes(x = r, y = obs)) +
  geom_ribbon(aes(ymin = lo, 
                  ymax = hi), fill = "gray80") +
  geom_line(col = "orange") +
  geom_line(aes(y = theo), color = "white") +
  xlab("Distance (km)") + ylab("K(r)") +
  labs(title = "Locations of lightning-caused wildfires in the ANF are consistent with a random process",
       subtitle = "Average observed (orange) versus modeled (white) number of additional wildfires within a distance r of a random wildfire.\nThe gray ribbon is the 99% uncertainy band based on 100 simulations from the random process model.") +
  theme_minimal()

```

Monthly occurrence (relative frequency) of lightning-caused wildfires in the ANF.
```{r}
df <- LightningFires.sf %>%
  st_drop_geometry() %>%
  mutate(MonthF = factor(month.name[Month], 
                         levels = rev(month.name), 
                         ordered = TRUE)) %>%
  group_by(MonthF, .drop = FALSE) %>%
  summarize(nF = n(),
            perF = nF/nrow(LightningFires.sf))

ggplot(data = df,
       mapping = aes(y = MonthF, 
                     x = perF,
                     fill = perF)) +
  geom_col() +
  scale_fill_distiller(palette = "Oranges",
                       direction = 1,
                       guide = FALSE) +
  scale_x_continuous(labels = percent) +
  xlab("") + ylab("") +
  labs(title = "Over 80% of lightning-caused wildfires in the Apalachicola National Forest occur during May-July",
       subtitle = "Based on data from 1992-2015",
       caption = "Data source: Short, Karen (2017)") +
  theme_dark() 
```

Annual number of lightning-caused fires and annual total amount of acres burned.
```{r}
df <- LightningFires.sf %>%
  st_drop_geometry() %>%
  filter(Month %in% c(5, 6, 7)) %>%
  mutate(YearF = factor(Year, levels = as.character(1992:2015), ordered = TRUE)) %>%
  group_by(YearF, .drop = FALSE) %>%
  summarize(nFires = n(),
            totalAcresBurned = sum(FIRE_SIZE)) %>%
  mutate(Year = 1992:2015)

ggplot(data = df,
       mapping = aes(x = Year, y = nFires, color = totalAcresBurned)) +
  geom_line(col = "gray80") +
  geom_point(size = 3) +
  scale_color_distiller(palette = "Oranges",
                        direction = 1,
                        guide = FALSE) +
  scale_x_continuous(breaks = seq(1992, 2015, by = 3)) +
  theme_minimal() +
  labs(x = "", y = "",
       title = "Large interannual variability in the number of lightning-caused wildfires in the ANF",
       subtitle = "Darker color points indicates more area burned")

ggplot(data = df,
       mapping = aes(x = Year, y = totalAcresBurned, color = nFires)) +
  geom_line(col = "gray80") +
  geom_point(size = 3) +
  scale_color_distiller(palette = "Oranges",
                        direction = 1,
                        guide = FALSE) +
  scale_x_continuous(breaks = seq(1992, 2015, by = 3)) +
  theme_minimal() +
  labs(x = "", y = "",
       title = "Large interannual variability in total acres burned from lightning-caused wildfires in the ANF",
       subtitle = "Darker color points indicates a larger number of fires")
```

Annual temporal autocorrelation in frequency and total area burned.
```{r}
plot(acf(df$nFires))
plot(acf(df$totalAcresBurned))
cor.test(df$nFires, df$totalAcresBurned)
cor.test(df$nFires[-1], df$totalAcresBurned[-24])
```

Possible project: Make fire size an ordered factor and create a model that predicts the probability by size category similar to EF rating.
```{r}
Fires.sf %>%
  group_by(FIRE_SIZE_) %>%
  summarize(nFires = n(),
            Range1 = range(FIRE_SIZE)[1],
            Range2 = range(FIRE_SIZE)[2])
```

https://www.nifc.gov/nicc/predictive/outlooks/outlooks.htm
https://www.cpc.ncep.noaa.gov/products/predictions/90day/

## Get TLH airport daily weather data and compute the Keetch & Byram drought index

Original paper outlining the rationale and how to create it: https://www.srs.fs.usda.gov/pubs/rp/rp_se038.pdf

The Keetch-Byram Drought Index assesses the risk of fire by representing the net effect of evapotranspiration and precipitation in producing cumulative moisture deficiency in deep duff and upper soil layers. The index ranges from zero, the point of no moisture deficiency, to 800, the maximum drought that is possible. Units are 100's of inches. KBDI relates current and recent weather conditions to potential or expected fire behavior. It was originally developed for the Southeast United States and is based primarily on recent rainfall patterns. It is one of the only drought indexes specifically developed to equate the effects of drought with potential fire behavior. 

For different soil types, the depth of soil required to hold 8 inches of moisture varies (loam = 30", clay = 25" and sand = 80"). A prolonged drought (high KBDI) influences fire intensity largely because more fuel is available for combustion (i.e. fuels have a lower moisture content). In addition, the drying of organic material in the soil can make it harder to suppress fires.

High values of the KBDI indicate conditions favorable for the occurrence and spread of wildfires, but drought is not by itself a prerequisite for wildfires. Other weather factors, such as wind, temperature, relative humidity and atmospheric stability, play a major role in determining the actual fire danger.

NWS Tallahassee daily weather data:
https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USW00093805/detail
https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/readme.txt
Column explanations: https://docs.google.com/document/d/1q2WEpXndpMx9lUq-0ON63GojkaOzixrGyYEhI_xkPtw/edit?usp=sharing
AWND = Average daily wind speed (tenths of meters per second) 

Import the summary of the day data and add columns to the data frame. Assign no rainfall on days with missing values.
```{r}
TLH.df <- read.csv(file = 'Data/TLH_Daily1940.csv',
                   stringsAsFactors = FALSE,
                   header = TRUE) %>%
  mutate(Date = as.Date(DATE)) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24 = replace_na(Rainfall24, 0),
         Rainfall24mm = Rainfall24 * 25.4)

TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 96 # missing max temperature. Filled from weather underground data

sum(is.na(TLH.df$PRCP))
nrow(TLH.df)
sum(is.na(TLH.df$PRCP)) / nrow(TLH.df) * 100  # only 5 missing values [< .02% of all days]
```

Compute daily values of the KBDI. Original paper outlining the rationale and how to create it: https://www.srs.fs.usda.gov/pubs/rp/rp_se038.pdf with a minor correction identified in Alexander1992.pdf (paper found in Dropbox/Literature).

Step one: Compute net rainfall on each day. Units are inches.
```{r}
Rainfall24 <- TLH.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0

CumR <- 0
NetR <- numeric()

for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH.df$NetR <- NetR
```

Step two: Compute daily drought index. Units of temperature are degrees F.
```{r}
Q <- 269  
R <- 59.23 # average annual rainfall for TLH

MaxTemp <- TLH.df$MaxTemp

Ql <- numeric()
DeltaQl <- numeric()
for(i in 1:length(Rainfall24)){
  DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
  Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
  Q <- ifelse(Q < 0, 0, Q) 
  Ql <- c(Ql, Q)
  DeltaQl <- c(DeltaQl, DeltaQ)
}

TLH.df$Ql <- Ql
TLH.df$Qlm <- Ql * .254  # tenth of an inch to mm
TLH.df$DeltaQl <- DeltaQl
TLH.df$DroughtIndex <- floor(Ql/100)

TLH.df <- TLH.df %>%
 dplyr::filter(Year >= 1946 & Year <= 2019) # Only full years
range(TLH.df$Ql)
```

Note: There is a package to compute the standaridized precip-evapotranspiration index {SPEI}. Only good for monthly data. Same for the Palmer Drought Severity Index (PDSI).

Monthly average soil moisture deficit.
```{r}
TLH.df %>%
  group_by(month) %>%
  summarise(Avg = mean(Qlm)) %>%
  ggplot(aes(x = month, y = Avg, fill = Avg)) +
  geom_col() +  
  scale_fill_gradient2(low = "#FC8B93", mid = "#965784", high = "#D9DE6E", guide = FALSE) +
  ylab("") + xlab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Monthly average soil moisture deficit (mm), Tallahassee, FL", 
       subtitle = "Amount of water needed to bring soil moisture to full capacity",
       caption = "Based on data from the NWSFO, Tallahassee, FL (1946-2019)") 
```

Make some trend plots. The second plot is statistically more realistic because it removes the temporal autocorrelations.
```{r}
ggplot(TLH.df, aes(x = Date, y = Qlm, color = Qlm)) +
  geom_line() +
  scale_color_gradient2(low = "#FC8B93", mid = "#965784", high = "#D9DE6E", guide = FALSE) +
#  scale_y_reverse(breaks = seq(0, 200, by = 50), labels = c("Wet", 50, 100, 150, "Dry")) +
  ylab("") + xlab("") +
  geom_smooth(method = lm, se = FALSE) +
  facet_wrap(~ month, ncol = 12) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Daily soil moisture deficit (mm), Tallahassee, FL", 
       subtitle = "Amount of water needed to bring soil moisture to full capacity",
       caption = "Based on data from the NWSFO, Tallahassee, FL (1946-2019)") 

TLH.df %>%
  group_by(month, Year) %>%
  summarise(Avg = mean(Qlm)) %>%
  ggplot(aes(x = Year, y = Avg, color = Avg)) +
  geom_point() +  
  scale_color_gradient2(low = "#FC8B93", mid = "#965784", high = "#D9DE6E", guide = FALSE) +
  ylab("") + xlab("") +
  geom_smooth(method = lm) +
  facet_wrap(~ month, ncol = 12) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Monthly average soil moisture deficit (mm) by year, Tallahassee, FL", 
       subtitle = "Amount of water needed to bring soil moisture to full capacity",
       caption = "Based on data from the NWSFO, Tallahassee, FL (1946-2019)") 
```

Side-by-side daily distribution plots (early vs late) by month. 1943:1980, 1981:2019
```{r}
TLH.df %>%
  mutate(Epoch = Year > 1980) %>%
  ggplot(aes(x = Qlm, color = Epoch)) +
#  geom_density(bw = 10, size = 1) +
  geom_freqpoly(size = 1, bins = 8) +
  scale_color_manual(labels = c("1943-1980", "1981-2019"), values = c("black", "sandybrown")) +
  scale_x_continuous(limits = c(0, 250), breaks = c(0, 100, 200)) +
  ylab("Number of days") + xlab("Daily soil moisture deficit (mm)") +
  facet_wrap(~ month, ncol = 12) +
#  coord_flip() +
  theme_minimal()
```

Repeat using data from Apalachacola (see `compare.rmd`). https://www.ncdc.noaa.gov/  > I want to search for data at a particular location > DATA TOOLS > Final a Station

Linear trend model.
```{r}
summary(lm(Qlm ~ Date, data = TLH.df)) # this assumes daily Qlm are independent (they are certainly not)

library(broom)
library(quantreg)

Trends.df <- TLH.df %>% 
  group_by(Year, month) %>%
  summarize(AvgQlm = mean(Qlm)) %>%
  group_by(month) %>%
#  do(tidy(rq(AvgQlm ~ Year, tau = .5, data = .))) %>%
  do(tidy(lm(AvgQlm ~ Year, data = .))) %>%
  dplyr::filter(term == "Year")

Avg <- TLH.df %>%
  group_by(month) %>%
  summarize(AvgQlm = mean(Qlm)) %>%
  pull(AvgQlm)

cor(Trends.df$estimate, Avg) # positive indicating upward trends are occurring in months with higher drought index values

Avg <- TLH.df %>%
  group_by(month) %>%
  summarize(nYear = n_distinct(Year), 
            TotalPRCP = sum(PRCP, na.rm = TRUE),
            AvgMonthlyPRCP = TotalPRCP/nYear) %>%
  pull(AvgMonthlyPRCP)

cor(Trends.df$estimate, Avg)  # negative as expected since upward trends are occurring in months with less rainfall
```

Soil moisture deficits are increasing at .4 mm/year on average during April. Which amounts to 4 mm/decade.

What is the correlation between PRCP, TMAX and Qlm grouped by month?
```{r}
TLH.df %>%
  group_by(month) %>%
  summarize(rPRCP = cor(Qlm, PRCP, use = "complete"),
            rTMAX = cor(Qlm, TMAX, use = "complete"),
            rTMIN = cor(Qlm, TMIN, use = "complete"),
            rTAVG = cor(Qlm, TAVG, use = "complete"))
```

Seasonal (fire season) temperature vs KDBI scatter plot. See TempVsPrecipSeasonal.jpeg
Create a character variable called `FireSeason` (May, June, July). Then group by that variable computing the average temp and total precipitation (or avg KDBI).
```{r}
df <- TLH.df %>%
  mutate(FireSeason = month %in% c("May", "Jun", "Jul")) %>%
  filter(FireSeason) %>%
  group_by(Year) %>%
  summarize(AvgHiTemp = mean(MaxTemp),
            AvgLoTemp = mean(MinTemp),
            TotalPrcp = sum(Rainfall24mm),
            AvgQlm = mean(Qlm))

ggplot(df, 
       mapping = aes(y = TotalPrcp, x = AvgHiTemp, label = Year, color = factor(Year))) +
  geom_point() +
  geom_label_repel(size = 2.5) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_color_grey(start = .8, end = .2, guide = FALSE) +
#  geom_smooth(method = lm, se = FALSE, col = "black") +
  ylab("Total Precipitation (mm)") + xlab("Average Daily High Temperature (°F)") +
  ggtitle(label = "Apalachicola National Forest fire seasons (May-Jul) that are drier tend to be hotter",
          subtitle = "Dry days are sunnier and thus they tend to get hotter. Later years have darker labels.") +
  theme_minimal() 
```

Get SOI data and join with TLH.df.
```{r}
SOI.df <- read.csv(file = "https://www.ncdc.noaa.gov/teleconnections/enso/indicators/soi/data.csv",
                   skip = 1, header = TRUE) %>%
  mutate(Date = parse_date_time(as.character(Date), "ym"),
         Year = year(Date),
         Month = month(Date),
         month = month(Date, label = TRUE, abbr = TRUE),
         SOI = Value) %>%
  dplyr::select(Year, Month, month, SOI)

Joined.df <- TLH.df %>% 
  dplyr::filter(Year >= 1951) %>%
  group_by(Year, month) %>%
  summarize(AvgQlm = mean(Qlm)) %>%
  dplyr::select(Year, month, AvgQlm) %>%
  left_join(SOI.df, by = c("Year", "month"))

Joined.df %>%
  group_by(month) %>%
  do(tidy(lm(AvgQlm ~ Year + SOI, data = .))) %>%
  dplyr::filter(term == "Year")

Joined.df %>%
  group_by(month) %>%
  do(tidy(lm(AvgQlm ~ Year + SOI, data = .))) %>%
  dplyr::filter(term == "SOI")
```

SOI is positively correlated with moisture deficit. La Nina conditions (positive SOI) during winter leads to larger moisture deficits (higher Qlm).

In terms of frequency by Drought Index category. Proportion of the year in each drought category.
```{r}
TLH.df %>%
#  dplyr::filter(month == "May") %>%
  group_by(Year, DroughtIndex) %>%
  summarize(N = n()) %>%
  ggplot(aes(x = Year, y = N)) +
  geom_col(aes(fill = as.factor(DroughtIndex))) +
  scale_fill_viridis_d(name = "KBDI") +
  scale_x_continuous(name = "", breaks = seq(1948, 2019, 4)) +
  scale_y_reverse() +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE))
```

Daily
```{r}
ggplot(TLH.df, aes(x = doy, y = Year, fill = Qlm)) +
  geom_tile() +
  scale_y_reverse(breaks = seq(1948, 2019, 4)) +
  scale_x_continuous(breaks = c(15,	46,	75,	106, 136, 167, 197, 228, 259, 289, 320, 350),
                     labels = month.abb, position = "top") +
  scale_fill_gradient2(name = "", low = "#FC8B93", mid = "#965784", high = "#D9DE6E") +
  ylab("") + xlab("") +
  theme_minimal() +
  ggtitle("Daily soil moisture deficit (mm), Tallahassee, FL", 
          subtitle = "Amount of water needed to bring soil moisture to full capacity") 

ggplot(TLH.df, aes(x = doy, y = Year, fill = factor(DroughtIndex))) +
  geom_tile() +
  scale_y_reverse(breaks = seq(1948, 2019, 4)) +
  scale_x_continuous(breaks = c(15,	46,	75,	106, 136, 167, 197, 228, 259, 289, 320, 350),
                     labels = month.abb, position = "top") +
  scale_fill_ordinal(name = "", direction = 1, alpha = .6) +
  ylab("") + xlab("") +
  theme_minimal() +
  guides(fill = guide_legend(reverse = TRUE)) +
  ggtitle("Daily drought category, Tallahassee, FL",
          subtitle = "7-6: Drought, 4-5: Dry, 1-3: Moist, 0: Saturated")
```

Number of days per year with KBDI above 650.
```{r}
library(MASS)

TLH.df %>%
  filter(Year != 2020) %>%
  group_by(Year) %>%
  summarize(nD = sum(Ql > 650)) %>%
  ggplot(aes(Year, nD)) +
  geom_point() +
  stat_smooth(method = "glm.nb",
              formula = y ~ x, 
              se = FALSE,
              col = "orange") 
```

Combine fires with KDBI.
```{r}
PSDI <- TLH.df %>%
  dplyr::filter(Year >= 1992 & Year <= 2015) %>%
  dplyr::filter(month == "Apr") %>%
  group_by(Year) %>%
  summarize(KBDI = last(Qlm))

SOI <- SOI.df %>%
  dplyr::filter(Year >= 1992 & Year <= 2015) %>%
  dplyr::filter(month == "Apr") %>%
  pull(SOI)

df <- data.frame(Year = PSDI$Year, nFires = Yearly$nF, KBDI = PSDI$KBDI, SOI = SOI)
```

Correlation by day of Month during April.
```{r}
TLH.df %>%
  dplyr::filter(Year >= 1992 & Year <= 2015) %>%
#  dplyr::filter(month == "Apr") %>%
#  mutate(Day = day(Date)) %>%
  group_by(doy, Year) %>%
  summarize(KBDI = Qlm) %>%
  dplyr::filter(doy != 366) %>%
  group_by(doy) %>%
  summarize(r = cor.test(KBDI, Yearly$nF)$estimate,
            rLo = cor.test(KBDI, Yearly$nF, conf.level = .9)$conf.int[1],
            rHi = cor.test(KBDI, Yearly$nF, conf.level = .9)$conf.int[2]) %>%
  ggplot(aes(x = doy, y = r)) +
  geom_ribbon(aes(ymin = rLo, ymax = rHi), fill = "gray90") +
  geom_point() +
  geom_line() +
  geom_segment(lineend = "round", aes(x = 121, xend = 213, y = .9, yend = .9), size = 3, col = "red") +
  scale_y_continuous(limits = c(-1, 1)) +
  scale_x_continuous(breaks = c(1,	32,	60,	91, 121, 152, 182, 213, 244, 274, 305, 335),
                     labels = month.abb, position = "top") +
  theme_minimal()
```

Negative binomial regression.
```{r}
library(MASS)

var(df$nFires)/mean(df$nFires)
exp(coef(glm.nb(nFires ~ 1, data = df)))
exp(coef(glm.nb(nFires ~ KBDI + SOI, data = df))) # 1.189 or 19% increase in the risk of a wildfire per cm increase in soil moisture deficit


model <- glm.nb(nFires ~ KBDI, data = df)

ggplot(df, aes(x = KBDI, y = nFires)) +
  geom_point() + 
  geom_line(aes(y = fitted(model))) +
  ylab("Number of Fires (May-July)") + xlab("Drought Index (April)") +
  theme_minimal()
```

Use the {brms} and {tidybayes} packages to get posterior distributions. See https://www.youtube.com/watch?v=wbzfqh_3LyM&list=PLCrWEzJgSUqzI3goQEAKkDsHg72inmqbe&index=20&t=0s

Also hot-days.Rmd, PredictTorFreqByEF.Rmd
