---
title: "Compare KDBI from TLH and AAF"
author: "Zach"
date: "12/11/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Here we compare the KDBI computed using data from TLH and from AAF.

The method to compute KDBI is in the _KBDI.Rmd_ file.

## Import the summary of the day data and add columns to the data frame

Assign no rainfall on days with missing values.

For Tallahassee Station TLH
```{r}
library(lubridate)
library(tidyverse)
TLH.df <- read_csv(file = 'Data/TLH_Daily1940.csv') %>%
  rename(Date = DATE) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24 = replace_na(Rainfall24, 0),
         Rainfall24mm = Rainfall24 * 25.4)
```

Fill in the missing temperature data.
```{r}
TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 96
```

Percent of missing precipitation days
```{r}
sum(is.na(TLH.df$PRCP))
nrow(TLH.df)
sum(is.na(TLH.df$PRCP)) / nrow(TLH.df) * 100
```

For Apalachicola Station KAAF
```{r}
KAAF.df <- read_csv(file = 'Data/KAAF_Daily1940.csv') %>%
  rename(Date = DATE) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24 = replace_na(Rainfall24, 0),
         Rainfall24mm = Rainfall24 * 25.4)
```

Filter KAAF.df to the longest matching period of no missing data
Oct. 9 2002 - Mar. 1 2004

```{r}
KAAFcompare.df <- KAAF.df %>%
  filter(KAAF.df$Date >= "2002-10-09" & KAAF.df$Date <= "2004-03-01")
```

Filter TLH.df to match time period of KAAFcompare.df

```{r}
TLHcompare.df <- TLH.df %>%
  filter(TLH.df$Date >= "2002-10-09" & TLH.df$Date <= "2004-03-01")
```




### Checking that there is no missing temp or precip data

For KAAFcompare
```{r}
sum(is.na(KAAFcompare.df$TMAX))
sum(is.na(KAAFcompare.df$PRCP))
nrow(KAAFcompare.df)
```

For TLHcompare
```{r}
sum(is.na(TLHcompare.df$TMAX))
sum(is.na(TLHcompare.df$PRCP))
nrow(TLHcompare.df)
```




## Compute daily values of the KBDI For TLHcompare

```{r}
Rainfall24 <- TLHcompare.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0
CumR <- 0
NetR <- numeric()
for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}
TLHcompare.df$NetR <- NetR
Q <- 269  
R <- 59.23
MaxTemp <- TLHcompare.df$MaxTemp
Ql <- numeric()
DeltaQl <- numeric()
for(i in 1:length(Rainfall24)){
  DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
  Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
  Q <- ifelse(Q < 0, 0, Q) 
  Ql <- c(Ql, Q)
  DeltaQl <- c(DeltaQl, DeltaQ)
}
TLHcompare.df$Ql <- Ql
TLHcompare.df$Qlm <- Ql * .254  # tenth of an inch to mm
TLHcompare.df$DeltaQl <- DeltaQl
TLHcompare.df$DroughtIndex <- floor(Ql/100)
range(TLHcompare.df$Ql)
```

### Compute daily values of the KBDI For KAAFcompare


```{r}
Rainfall24 <- KAAFcompare.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0
CumR <- 0
NetR <- numeric()
for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}
KAAFcompare.df$NetR <- NetR
Q <- 269  
R <- 59.23
MaxTemp <- KAAFcompare.df$MaxTemp
Ql <- numeric()
DeltaQl <- numeric()
for(i in 1:length(Rainfall24)){
  DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
  Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
  Q <- ifelse(Q < 0, 0, Q) 
  Ql <- c(Ql, Q)
  DeltaQl <- c(DeltaQl, DeltaQ)
}
KAAFcompare.df$Ql <- Ql
KAAFcompare.df$Qlm <- Ql * .254  # tenth of an inch to mm
KAAFcompare.df$DeltaQl <- DeltaQl
KAAFcompare.df$DroughtIndex <- floor(Ql/100)
range(KAAFcompare.df$Ql)
```



###Start Comparing KAAF and TLH Daily KBDI values


Perform left join to merge data frames.

.x = KAAF Station Data
.y = TLH Station Data

*note in ByMonth, 2003 is the only full year

```{r}
library(tidyverse)
KAAFcompare.df <- KAAFcompare.df %>%
  dplyr::select(NAME, Date, MaxTemp, Rainfall24mm, NetR, Qlm)
TLHcompare.df <- TLHcompare.df %>%
  dplyr::select(NAME, Date, MaxTemp, Rainfall24mm, NetR, Qlm)
CompareLong.df <- rbind(KAAFcompare.df, TLHcompare.df)
CompareWide.df <- left_join(KAAFcompare.df, TLHcompare.df, by = "Date") %>%
  mutate(Year = year(Date), 
         Month = month(Date))
ByYear <- CompareWide.df %>%
  group_by(Year) %>%
  summarise(cc = cor(Qlm.x, Qlm.y))
ByMonth <- CompareWide.df %>%
  group_by(Month) %>%
  summarise(cc = cor(Qlm.x, Qlm.y))
```


Make plots
```{r}
library(ggplot2)
ggplot(data = CompareLong.df, aes(x = Date, y = Qlm, color = NAME)) +
  geom_line() 
```



###Plots comparing Ql for TLH and APL

1999 is the year on record with the lowest correlation coefficient. 

```{r}
library(ggplot2)
joineddata.df %>%
  filter(Year.x == '1999') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  ylab("KBDI")
  #legend("right",
  #c("KAAF KDBI", "TLH KBDI"),
  #col = c("darkred", "steelblue"))
#testing file
```

How do the cc values for 1999 vary over each month?

```{r}
joineddata.df %>%
  filter(Year.x == '1999') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  facet_wrap(~ month.x, scales = "free_x") +
  ylab("KBDI")
```

1998 is the year with the highest cc value.
```{r}
joineddata.df %>%
  filter(Year.x == '1998') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  ylab("KBDI")
```

How do the values for cc vary over each month in 1998.

```{r}
joineddata.df %>%
  filter(Year.x == '1998') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  facet_wrap(~ month.x, scales = "free_x") +
  ylab("KBDI")
```
Our plots show that July is where the Ql values at each station vary the most.
Plot the Ql values for July across every year to check if this trend is consistent

```{r}
joineddata.df %>%
  filter(month.x == 'Jul') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  facet_wrap(~ Year.x, scales = "free_x") +
  ylab("KBDI")
```
Break time up into seasons.Plot a 6 month dry season and a 6 month wet season. 
Dry period: May - Oct

```{r}
joineddata.df %>%
  filter(month.x == 'May' | month.x == 'Jun' | month.x == 'Jul' | month.x == 'Aug' | month.x == 'Sep.' | month.x == 'Oct') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  facet_wrap(~ Year.x, scales = "free_x") +
  ylab("KBDI")
```
Further Filtering summer dry period: Taking out May and October

```{r}
joineddata.df %>%
  filter(month.x == 'Jun' | month.x == 'Jul' | month.x == 'Aug' | month.x == 'Sep') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  facet_wrap(~ Year.x, scales = "free_x") +
  ylab("KBDI")
```

Wet Period: November - April
```{r}
joineddata.df %>%
  filter(month.x == 'Nov' | month.x == 'Dec' | month.x == 'Jan' | month.x == 'Feb' | month.x == 'Mar' | month.x == 'Apr') %>%
ggplot(aes(x = Date)) +
  geom_line(aes(y = Ql.x), color = "darkred") +
  geom_line(aes(y = Ql.y), color = "steelblue") +
  facet_wrap(~ Year.x, scales = "free_x") +
  ylab("KBDI")
```

How many more rainy days is KAAF having compared to TLH in July?

```{r}
joineddata.df %>%
  filter(month.x == 'July') %>%
  group_by(Year.x) %>%
  summarise(Nraindays = sum(Rainfall24mm.x > 0)) %>%
ggplot(aes(x = Year.x, y = Nraindays, fill = Nraindays)) +
  geom_histogram()
```





###Notes/ Extra Code Chunks

This code chunk is for notes purposes at the moment 

```{r}
library(lubridate)
library(dplyr)
library(tidyr)
KAAF.df <- read.csv(file = 'Data/KAAF_Daily1940.csv',
                   stringsAsFactors = FALSE,
                   header = TRUE) %>%
  mutate(Date = as.Date(DATE)) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24 = replace_na(Rainfall24, 0),
         Rainfall24mm = Rainfall24 * 25.4)
#Missing max temperatures. Data not found on weather underground. Filled with normals from nws: https://www.weather.gov/tae/apalachicola_normalsrecords
# x <- c(1, 2, 4, NA, 6, NA, 10)
# is.na(x)
# y <- c(20, 30)
# x[is.na(x)] <- y
# x
#updated code filling missing max temp data.
x <- c(KAAF.df$MaxTemp)
is.na(x)
y <- c(88,88,88,76,76,73,63,66,80,78,72,72,72,73,73,73,72,78,77,70,64,63,62,72,80,69,74,74,75,91,89,88,86,87,88,87,82,84,87,87,87,85,81,84,84,82,86,86,54,62,54,50,75,82,87,89,91,88,89,89,81,73,61,60,54,56,64,63,75,80,88,87,90,88,84,79,78,54,65,72,51,64,72,68,69,64,61,64,73,68,72,74,75,84,82,91,88,93,90,91,87,88,86,88,84,85,86,77,81,82,82,79,83,75,77,75,71,75,65,55,74,72,63,70,77,71,66,59,68,68,74,72,73,75,77,65,67,76,78,78,76,66,87,78,77,85,81,81,79,83,87,84,84,89,91,88,88,88,90,88,83,86,85,86,84,85,88,91,90,88,84,82,90,83,85,79,84,84,80,79,84,83,87,83,82,8372,69,74,72,62,74,70,70,72,50,57,64,57,65,64,62,70,64,56,79,73,75,74,69,72,72,65,72,81,80,78,78,89,88,89,86,86,84,89,67,78,78,72,74,72,71,74,66,68,71,69,75,72,67,62,65,62,64,68,72,65,70,64,49,50,62,54,61,63,66,66,56,57,54,55,57,68,58,54,56,56,40,55,63,63,46,63,69,52,53,67,70,64,67,66,71,72,76,77,81,81,89,85,88,92,91,83,91,90,86,91,91,84,87,83,85,87,85,82,83,82,84,79,82,78,74,70,75,76,64,59,44,71)
x[is.na(x)] <- y
x
# 
# KAAF.df$MaxTemp[KAAF.df$Date == "1964-07-09"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "1964-07-10"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "1964-07-11"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "1964-11-03"] <- 76
# KAAF.df$MaxTemp[KAAF.df$Date == "1964-11-05"] <- 76
# KAAF.df$MaxTemp[KAAF.df$Date == "1996-11-15"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "1996-12-31"] <- 63 #7 degrees lower than day before and after. 12/30 +  1/1 maxTemp = 70.
# KAAF.df$MaxTemp[KAAF.df$Date == "1997-02-14"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "1997-04-30"] <- 80 #4/29 = maxTemp = 74, 5/1 maxTemp = 78. Normal of 80 does not fit in trend? 
# KAAF.df$MaxTemp[KAAF.df$Date == "1997-10-29"] <- 78 #strong cold front on 10/27. Normal of 78 does not fit trend
# KAAF.df$MaxTemp[KAAF.df$Date == "2000-03-25"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2000-03-26"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2000-03-27"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2000-03-28"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "2000-03-29"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "2000-03-30"] <- 73
# 
# #Missing max temperatures. Filled with Weather Underground data.
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-01"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-02"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-03"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-04"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-05"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-10"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-11"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-12"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-13"] <- 80
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-14"] <- 69
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-15"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-16"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-03-31"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-08-31"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-09-01"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-09-02"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-09-03"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-09-04"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-09-05"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2001-09-06"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-06"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-07"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-08"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-09"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-10"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-11"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-25"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-26"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-27"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-05-28"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-10-07"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2002-10-08"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-01-31"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-03-01"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-03-02"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-03-03"] <- 50
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-06-01"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-06-02"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-06-30"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-07-28"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-09-02"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-10-01"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-10-02"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-10-04"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-10-18"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-12-04"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-12-09"] <- 61
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-12-10"] <- 60
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-12-11"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2005-12-21"] <- 56
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-01-31"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-03-02"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-04-12"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-04-16"] <- 80
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-05-29"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-06-04"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-06-09"] <- 90
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-09-04"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-10-03"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-10-16"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-10-27"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-10-28"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-02"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-09"] <- 59
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-12"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-17"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-27"] <- 51
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-28"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-29"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2006-12-30"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-01-05"] <- 69
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-01-09"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-01-18"] <- 61
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-02-20"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-02-21"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-03-07"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-03-08"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-03-30"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-03-31"] <- 75
# 
# #All of April 2007 Data Missing
# 
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-05-11"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-05-28"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-06-10"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-06-18"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-07-01"] <- 93
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-07-02"] <- 90
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-07-31"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-09-03"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-09-08"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-09-14"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-09-27"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-09"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-10"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-11"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-12"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-13"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-14"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-15"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-16"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-17"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-10-19"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-11-21"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-11-22"] <- 75
# # KAAF.df$MaxTemp[KAAF.df$Date == "2007-11-25"] <- missing data on weather underground -normal = 71
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-11-26"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-11-27"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-12-17"] <- 55
# KAAF.df$MaxTemp[KAAF.df$Date == "2007-12-30"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-02-17"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-02-19"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-02-21"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-02-26"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-04"] <- 71
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-07"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-08"] <- 59
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-11"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-13"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-15"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-17"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-18"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-19"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-23"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-24"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-03-31"] <- 67
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-02"] <- 76
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-04"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-05"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-08"] <- 76
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-15"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-22"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-04-29"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-02"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-05"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-07"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-12"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-14"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-25"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-29"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-30"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-05-31"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-01"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-02"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-03"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-04"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-05"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-06"] <- 90
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-07"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-22"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-06-30"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-01"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-06"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-07"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-08"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-12"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-20"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-22"] <- 90
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-07-28"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-08-17"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-08-23"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-09-06"] <- 90
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-09-24"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-09-28"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-02"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-04"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-08"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-09"] <- 80
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-10"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-11"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-12"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-13"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-14"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-15"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-17"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-27"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-10-30"] <- 69
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-11-07"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-11-11"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-11-23"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-11-28"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2008-11-30"] <- 70
# #Excel file missing all data from Dec 2008
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-01-03"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-01-05"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-01-17"] <- 50
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-01-22"] <- 57
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-01-31"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-02-06"] <- 57
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-02-08"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-02-14"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-02-17"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-02-19"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-02-22"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-02"] <- 56
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-09"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-10"] <- 73
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-19"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-24"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-29"] <- 69
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-03-31"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-01"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-08"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-16"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-23"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-24"] <- 80
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-26"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-04-30"] <- 78
# #Excel Missing data: May 2009 - July 2009
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-08-05"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-08-18"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-09-09"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-09-14"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-09-20"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-10-04"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-10-12"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-10-19"] <- 67
# #October 26th data missing from Weather Underground - normal = 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-08"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-09"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-14"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-15"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-16"] <- 71
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-17"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-18"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-19"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-20"] <- 71
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-21"] <- 69
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-22"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-23"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-24"] <- 67
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-25"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-26"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-27"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-28"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-29"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-11-30"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-01"] <- 65
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-02"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-03"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-04"] <- 49
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-11"] <- 50
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-17"] <- 62
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-21"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-22"] <- 61
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-23"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-24"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-25"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-26"] <- 56
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-27"] <- 57
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-28"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-29"] <- 55
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-30"] <- 57
# KAAF.df$MaxTemp[KAAF.df$Date == "2009-12-31"] <- 68
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-01"] <- 58
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-02"] <- 54
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-07"] <- 56
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-08"] <- 56
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-09"] <- 40
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-14"] <- 55
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-20"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-02-02"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-02-12"] <- 46
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-02-20"] <- 63
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-02-23"] <- 69
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-02-24"] <- 52
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-02-25"] <- 53
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-03-10"] <- 67
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-03-19"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-03-21"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-03-24"] <- 67
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-03-25"] <- 66
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-01-31"] <- 71
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-04-02"] <- 72
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-04-07"] <- 76
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-04-27"] <- 77
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-05-05"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-05-11"] <- 81
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-05-18"] <- 89
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-05-25"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-05-31"] <- 88
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-06-18"] <- 92
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-06-23"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-07-02"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-07-10"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-07-13"] <- 90
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-07-14"] <- 86
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-08-04"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-08-17"] <- 91
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-08-30"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-09-23"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-07"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-08"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-09"] <- 87
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-10"] <- 85
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-11"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-12"] <- 83
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-13"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-14"] <- 84
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-15"] <- 79
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-10-25"] <- 82
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-11-01"] <- 78
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-11-15"] <- 74
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-11-17"] <- 70
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-11-22"] <- 75
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-11-23"] <- 76
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-12-16"] <- 64
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-12-25"] <- 59
# KAAF.df$MaxTemp[KAAF.df$Date == "2010-12-26"] <- 44
# KAAF.df$MaxTemp[KAAF.df$Date == "2011-10-29"] <- 71
# #Excel missing all data 4/11/2012 - 6/1/2012
sum(is.na(KAAF.df$PRCP))
nrow(KAAF.df)
sum(is.na(KAAF.df$PRCP)) / nrow(KAAF.df) * 100  #105 missing precip days
```

Fill the missing Apalacacola maximum temperature data

Dates missing dates before March 1st 2001 are filled with normals from the nws:
https://www.weather.gov/tae/apalachicola_normalsrecords

Starting on March 1, 2001 all missing temperature values are filled from weather underground archives
*note:
10/26/2009 weather underground archive is missing. Value filled from normals = 78
11/25/2007 weather underground archive is missing. Value filled from normals = 71

Chunks of data are missing from the KAAF excel file.
No data:
All of April 2007
All of December 2008
May 2009 - July 2009
April 11, 2012 - June 1, 2012

```{r}
AllTemps <- KAAF.df$MaxTemp
#is.na(AllTemps)
OrderMissingT <- c(88,88,88,76,76,73,63,66,80,78,72,72,72,73,73,73,72,78,77,70,64,63,62,72,80,69,74,74,75,91,89,88,86,87,88,87,82,84,87,87,87,85,81,84,84,82,86,86,54,62,54,50,75,82,87,89,91,88,89,89,81,73,61,60,54,56,64,63,75,80,88,87,90,88,84,79,78,54,65,72,51,64,72,68,69,64,61,64,73,68,72,74,75,84,82,91,88,93,90,91,87,88,86,88,84,85,86,77,81,82,82,79,83,75,77,75,71,75,65,55,74,72,63,70,77,71,66,59,68,68,74,72,73,75,77,65,67,76,78,78,76,66,87,78,77,85,81,81,79,83,87,84,84,89,91,88,88,88,90,88,83,86,85,86,84,85,88,91,90,88,84,82,90,83,85,79,84,84,80,79,84,83,87,83,82,8372,69,74,72,62,74,70,70,72,50,57,64,57,65,64,62,70,64,56,79,73,75,74,69,72,72,65,72,81,80,78,78,89,88,89,86,86,84,89,67,78,78,72,74,72,71,74,66,68,71,69,75,72,67,62,65,62,64,68,72,65,70,64,49,50,62,54,61,63,66,66,56,57,54,55,57,68,58,54,56,56,40,55,63,63,46,63,69,52,53,67,70,64,67,66,71,72,76,77,81,81,89,85,88,92,91,83,91,90,86,91,91,84,87,83,85,87,85,82,83,82,84,79,82,78,74,70,75,76,64,59,44,71)
AllTemps[is.na(AllTemps)] <- OrderMissingT
KAAF.df$MaxTemp <- AllTemps
```

Comparing values


```{r}
joineddata.df <- left_join(KAAFcompare.df, TLHcompare.df, by = "Date")
joineddata.df %>% dplyr::select(Date, NAME.x, DroughtIndex.x, NAME.y, DroughtIndex.y)
cor(KAAFcompare.df$DroughtIndex,TLHcompare.df$DroughtIndex)
cor(KAAFcompare.df$Ql,TLHcompare.df$Ql)
joineddata.df %>%
  group_by(month.x) %>%
  summarise(cc = cor(Ql.x,Ql.y))
joineddata.df %>%
  group_by(Year.x) %>%
  summarise(cc = cor(Ql.x,Ql.y))
```
```{r}
joineddata.df %>%
  filter(month.x =='Apr')%>%
  group_by(Year.x) %>%
  summarise(cc = cor(Ql.x,Ql.y)) %>%
  pull(cc)
```




### Trends

Make some trend plots. The second plot is statistically more realistic because it removes the temporal autocorrelations.
```{r}
library(ggplot2)
ggplot(KAAF.df, aes(x = Date, y = Qlm, color = Qlm)) +
  geom_line() +
  scale_color_gradient2(low = "#FC8B93", mid = "#965784", high = "#D9DE6E", guide = FALSE) +
#  scale_y_reverse(breaks = seq(0, 200, by = 50), labels = c("Wet", 50, 100, 150, "Dry")) +
  ylab("") + xlab("") +
  geom_smooth(method = lm, se = FALSE) +
  facet_wrap(~ month, ncol = 12) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Daily soil moisture deficit (mm), Apalachicola, FL", 
       subtitle = "Amount of water needed to bring soil moisture to full capacity",
       caption = "Based on data from the NWSFO, Apalachicola, FL (1940-2012)") 
KAAF.df %>%
  dplyr::group_by(month, Year) %>%
  dplyr::summarise(Avg = mean(Qlm)) %>%
  ggplot(aes(x = Year, y = Avg, color = Avg)) +
  geom_point() +  
  scale_color_gradient2(low = "#FC8B93", mid = "#965784", high = "#D9DE6E", guide = FALSE) +
  ylab("") + xlab("") +
  geom_smooth(method = lm) +
  facet_wrap(~ month, ncol = 12) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Monthly average soil moisture deficit (mm) by year, Apalachicola, FL", 
       subtitle = "Amount of water needed to bring soil moisture to full capacity",
       caption = "Based on data from the NWSFO, Apalachicola, FL (1940-2012)") 
```

Side-by-side daily distribution plots (early vs late) by month. 1940:1980, 1981:2012
Temperature distributions by epoch for days defined by the hot day/night seasons.
```{r}
KAAF.df %>%
  mutate(Epoch = Year > 1980) %>%
  ggplot(aes(x = Qlm, color = Epoch)) +
#  geom_density(bw = 10, size = 1) +
  geom_freqpoly(size = 1, bins = 8) +
  scale_color_manual(labels = c("1940-1980", "1981-2012"), values = c("black", "sandybrown")) +
  scale_x_continuous(limits = c(0, 250), breaks = c(0, 100, 200)) +
  ylab("Number of days") + xlab("Daily soil moisture deficit (mm)") +
  facet_wrap(~ month, ncol = 12) +
#  coord_flip() +
  theme_minimal()
```