---
title: "Climate Trends in Day-Of-The-Year High and Low Temperatures"
output: html_document
editor_options: 
  chunk_output_type: console
---

The upward trend is 0.12°C per decade from 1951-2012. The Intergovernmental Panel on Climate Change (IPCC) in their 2013 Fifth Assessment Report state “it is extremely likely that more than half of the observed global trend was caused by anthropogenic (human-activity) increase in greenhouse gas concentrations and other anthropogenic forcings”. 

1. Writing is thinking, 
2. 1 article, 1 question, 1 finding, 
3. it's not a review of the literature, it's an argument about why we need your paper

What day(s) of the year show the largest long-term (climate) trends in max/min temperatures? Corollary: what day(s) of the year do not show a normal distribution in max/min temperatures?

R version 4.1.0 (2021-05-18) -- "Camp Pontanezen"

## Get the required packages
```{r}
library(tidyverse)
library(lubridate)
library(broom)
```

## Get TLH airport daily weather data

NWS Tallahassee daily weather data:
https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USW00093805/detail
https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/readme.txt
Column explanations: https://docs.google.com/document/d/1q2WEpXndpMx9lUq-0ON63GojkaOzixrGyYEhI_xkPtw/edit?usp=sharing
AWND = Average daily wind speed (tenths of meters per second) 

Import the summary-of-the-day data and add columns to the data frame.
```{r}
TLH.df <- read.csv(file = 'Data/TLH_Daily1940-2020.csv',
                   stringsAsFactors = FALSE,
                   header = TRUE) %>%
  mutate(Date = as.Date(DATE)) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN) %>%
  filter(Year <= 2020 & Year >= 1943)

# Missing temperatures filled in from NCDC daily summaries https://gis.ncdc.noaa.gov/maps/ncei/summaries/daily

TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 95
TLH.df$MinTemp[TLH.df$Date == "2005-07-08"] <- 72
TLH.df$MinTemp[TLH.df$Date == "2002-10-08"] <- 67 # Quincy

TLH.df <- TLH.df |>
  mutate(AvgTemp = (MaxTemp + MinTemp) / 2)

( AnnualMeanTemperatures <- TLH.df %>%
  group_by(Year) %>%
  summarize(AvgHighTemp = mean(MaxTemp),
            AvgLowTemp = mean(MinTemp),
            AvgMeanTemp = mean(AvgTemp)) )
```

Plot annual mean temperatures.
```{r}
AnnualMeanTemperatures |>
  pivot_longer(cols = c(AvgHighTemp, AvgLowTemp)) |>
  ggplot(mapping = aes(x = Year, y = value, color = name)) +
  scale_y_continuous(limits = c(50, 90)) +
  scale_color_discrete(guide = "none") +
    geom_point() +
    geom_smooth() +
  theme_minimal() +
  labs(title = "Annual mean daily high and low temperatures (°F)",
       subtitle = "Tallahasee, Florida (1943-2020)",
       x = "", y = "")
```

Histogram of daily high temperature.
```{r}
labels <- c(paste0(seq(30, 100, by = 10), "°", " F"))

p1 <- TLH.df |> 
ggplot(mapping = aes(x = MaxTemp)) + 
  geom_histogram(binwidth = 1, 
                 mapping = aes(fill = ..count..)) +
  scale_fill_viridis_c(guide = "none") +
  scale_x_continuous(breaks = seq(30, 100, by = 10),
                     labels = labels) +
  theme_minimal() +
  labs(title = "Frequency of daily high temperatures",
       subtitle = "Tallahassee, Florida (1943-2020)",
       x = "", y = "")

arrows <- 
  tibble(
    x1 = c(75),
    x2 = c(89),
    y1 = c(1025), 
    y2 = c(1200)
  )

p1 +
  annotate("text", x = 60, y = 1000, label = "The most common high temperature is 90° F") +
  geom_curve(
    data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(.08, "inch")), size = .5,
    color = "gray20", curvature = -.3)
```

Non-normality in daily high and low temperatures
```{r}
library(moments)
library(nortest)

TLH.df |>
  group_by(doy) |>
  summarize( AvgTempByDoY = mean(AvgTemp),
             SDTempByDoY = sd(AvgTemp),
             SkewTempByDoY = skewness(AvgTemp) ) |>
  ggplot(mapping = aes(x = doy, y = SkewTempByDoY)) +
    geom_point() +
    geom_smooth()

normTest <- 
  TLH.df |>
#  filter(Year >= 1990) |>
  group_by(doy) |>
  summarize(pvalueS = shapiro.test(AvgTemp)$p.value,
            pvalueAD = ad.test(AvgTemp)$p.value) 

normTest$doy[normTest$pvalueS > .05]
normTest$doy[normTest$pvalueAD > .05]
```

Compute long-term trends in day-of-year high and low temperatures.
```{r}
( Trends <- 
  TLH.df |> 
    mutate(MaxTemp = (MaxTemp - 32) * 5/9) |>
#  filter(Year >= 1990) |>
    group_by(doy) |>
    do(tidy(lm(MaxTemp ~ Year, data = .))) |>
    filter(term == "Year") |>
    mutate(Date = as.Date(doy - 1, origin = "2020-01-01"),
           estimate = estimate * 10,
           std.error = std.error * 10,
           ymax = estimate + std.error,
           ymin = estimate - std.error) )

Trends |> 
  arrange(desc(statistic))

TLH.df |>
  filter(doy %in% c(26, 350, 192)) |>
  mutate(Day = factor(case_when(Day = doy == 26 ~ "26 January",
                         Day = doy == 350 ~ "15/16 December",
                         Day = doy == 192 ~ "10/11 July"), 
                      levels = c("26 January", "10/11 July", "15/16 December"))) |>
  select(Year, MaxTemp, Date, doy, Day) |>
  ggplot(mapping = aes(x = Year, y = MaxTemp)) +
    geom_point() +
    geom_smooth(method = lm) +
  facet_wrap(~ Day) +
  theme_minimal()
```

Plot trends.
```{r}
p1 <- Trends |>
  ggplot(mapping = aes(x = doy, y = estimate, color = statistic) ) +
  scale_y_continuous(limits = c(-1.2, 1.2)) +
  geom_hline(yintercept = 0, color = "gray70") +
  scale_color_distiller(palette = "RdBu", 
                        limits = c(-1, 1) * max(abs(Trends$statistic)),
                        guide = 'none') +
  geom_point() + 
  geom_errorbar(mapping = aes(ymin = ymin, ymax = ymax)) +
  scale_x_continuous(position = "bottom", 
                     breaks = c(1, 32, 61, 92, 122, 153, 183, 214, 245, 275, 306, 336) + 13,
                     labels = month.name) +
  theme_minimal() +
  labs(title = "Long-term (1943-2020) trends in daily high temperature (°C/decade) in Tallahassee, Florida", 
       subtitle = "Darker colors indicate greater evidence against a no-change hypothesis", 
       x = "", y = "")
```

Annotate plot. http://jenrichmond.rbind.io/post/idhtg-how-to-annotate-plots/

```{r}
  annotate("text", x = 46, y = -1, label = "Most cooling day: January 26") +
  annotate("text", x = 130, y = .8, label = "Most significant warming day: July 10 or 11") +
  annotate("text", x = 280, y = 1, label = "Most warming day: December 15 or 16") 
  
arrows <- 
  tibble(
    x1 = c(187, 212),
    x2 = c(205, 204),
    y1 = c(6000, 3000), 
    y2 = c(5500, 3500)
  )

arrows

p1 +
  geom_curve(
    data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.08, "inch")), size = 0.5,
    color = "gray20", curvature = -0.3)
```

Use `ggdist` package from Matthew Kay.
```{r}
library(ggdist)
# theme_set(theme_ggdist())

TLH.df |>
  ggplot(mapping = aes(x = doy, y = MaxTemp)) +
  stat_gradientinterval()

```