---
title: "Climate Trends in Daily Temperatures"
output: html_document
editor_options: 
  chunk_output_type: console
---

1. Writing is thinking, 
2. 1 article, 1 question, 1 finding, 
3. it's not a review of the literature, it's an argument about why we need your paper

What day(s) of the year show the largest long-term trends in max/min temperatures? Corollary: what day(s) of the year do not show a normal distribution in max/min temperatures?

R version 4.1.0 (2021-05-18) -- "Camp Pontanezen"

## Get the required packages
```{r}
library(tidyverse)
library(lubridate)
```

## Get TLH airport daily weather data

NWS Tallahassee daily weather data:
https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USW00093805/detail
https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/readme.txt
Column explanations: https://docs.google.com/document/d/1q2WEpXndpMx9lUq-0ON63GojkaOzixrGyYEhI_xkPtw/edit?usp=sharing
AWND = Average daily wind speed (tenths of meters per second) 

Import the summary of the day data and add columns to the data frame.
```{r}
TLH.df <- read.csv(file = 'Data/TLH_Daily1940-2020.csv',
                   stringsAsFactors = FALSE,
                   header = TRUE) %>%
  mutate(Date = as.Date(DATE)) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24mm = Rainfall24 * 25.4) %>%
  filter(Year <= 2020)

sum(is.na(TLH.df$PRCP))
sum(is.na(TLH.df$PRCP)) / nrow(TLH.df) * 100  # only 5 missing values [~ .017% of all days]
sum(is.na(TLH.df$TMAX))
sum(is.na(TLH.df$TMAX)) / nrow(TLH.df) * 100  # only 1 missing value [~ .0034% of all days]

# Missing temperature and rainfall filled in from NCDC daily summaries
# https://gis.ncdc.noaa.gov/maps/ncei/summaries/daily

TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 95
TLH.df$MinTemp[TLH.df$Date == "2005-07-08"] <- 72
TLH.df$MinTemp[TLH.df$Date == "2002-10-08"] <- 67 # Quincy


( AnnualMeanTemperatures <- TLH.df %>%
  filter(Year >= 1943) %>%  # complete data started in March 1943
  group_by(Year) %>%
  summarize(AvgHighTemp = mean(MaxTemp),
            AvgLowTemp = mean(MinTemp)) )
```

Plot annual mean temperatures.
```{r}
AnnualMeanTemperatures |>
  pivot_longer(cols = c(AvgHighTemp, AvgLowTemp)) |>
  ggplot(mapping = aes(x = Year, y = value, color = name)) +
    geom_point() +
    geom_smooth(method = lm)
```

Non-normality in daily high and low temperatures
```{r}
library(moments)
library(nortest)

TLH.df |>
  group_by(doy) |>
  summarize( AvgTempByDoY = mean(MinTemp),
             SDTempByDoY = sd(MinTemp),
             SkewTempByDoY = skewness(MinTemp) ) |>
  ggplot(mapping = aes(x = doy, y = SDTempByDoY)) +
    geom_point() +
    geom_smooth()

normTest <- 
  TLH.df |>
  filter(Year >= 1990) |>
  mutate(AvgTemp = (MaxTemp + MinTemp) / 2) |>
  group_by(doy) |>
  summarize(pvalueS = shapiro.test(MaxTemp)$p.value,
            pvalueAD = ad.test(MaxTemp)$p.value) 

normTest$doy[normTest$pvalueS > .05]
normTest$doy[normTest$pvalueAD > .05]
  
library(broom)
( Trends <- 
  TLH.df |> 
  filter(Year >= 1990) |>
  group_by(doy) |>
  do(tidy(lm(MinTemp ~ Year, data = .))) |>
  dplyr::filter(term == "Year") )
  
Trends |>
  ggplot(mapping = aes(x = doy, y = statistic)) +
  geom_point() + 
  geom_smooth()
```

Use `ggdist` package from Matthew Kay.
```{r}
library(ggdist)
theme_set(theme_ggdist())

TLH.df |>
  ggplot(mapping = aes(x = doy, y = MaxTemp)) +
  stat_gradientinterval()

```

Compute trends using `do()`.