---
title: "ExploratoryAnalysis_Committee"
author: "Zach"
date: "5/27/2021"
output: html_document
---

This file will explore ideas proposed by the committee during the prospectus defense on 5/25/2021


### Checking Value of Q

It is expected that with time KBDI plots will become equal regardless of the initial value of Q


Copy and paste KBDI code

## Import the summary of the day data and add columns to the data frame

Assign no rainfall on days with missing values.

For Tallahassee Station TLH
```{r}
library(lubridate)
library(tidyverse)
TLH.df <- read_csv(file = 'Data/TLH_Daily1940.csv') %>%
  rename(Date = DATE) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24 = replace_na(Rainfall24, 0),
         Rainfall24mm = Rainfall24 * 25.4)
```

Fill in the missing temperature data.
```{r}
TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 96
```


A new data frame is created to compare initial Q values. Q is representative of the starting value of KBDI. Here a column of several Q values is created to be used and compared among multiple plots. Daily KBDI is calculated for each initial Q value. This is testing the theory that with enough time elapsed, the initial value of Q does not matter.


```{r}
Rainfall24 <- TLH.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0
CumR <- 0
NetR <- numeric()
for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH.df$NetR <- NetR
R <- 59.23

rangeQ.df <- TLH.df %>%
  select(Date, Rainfall24, MaxTemp, NetR)

for(i in 1:9){
  Q <- i*100
  MaxTemp <- TLH.df$MaxTemp
  Ql <- numeric()
  for(i in 1:length(Rainfall24)){
    DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
    Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
    Q <- ifelse(Q < 0, 0, Q) 
    Ql <- c(Ql, Q)
  }
  rangeQ.df[, ncol(rangeQ.df)+1] <- Ql
  colnames(rangeQ.df) <- c("Date", "Rainfall24", "MaxTemp", "NetR", "Q0", "Q100", "Q200", "Q300", "Q400", "Q500", "Q600", "Q700", "Q800")
}
```


```{r}
library(lubridate)
library(ggplot2)

rangeQ.df %>% 
  filter(year(Date) == 1940) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(aes(y = Q0), colour = "red") +
    geom_line(aes(y = Q400), colour = "blue") + 
    geom_line(aes(y = Q800))
``` 



```{r}
library(lubridate)
library(ggplot2)

rangeQ.df %>% 
  filter(year(Date) == 1940) %>%
  filter(month(Date) == '6' | month(Date) == '7' | month(Date) == '8') %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(aes(y = Q0), colour = "red") +
    geom_line(aes(y = Q400), colour = "blue") + 
    geom_line(aes(y = Q800))
``` 






##Explore lightning data and concerns mentioned by committee.

Get lightning data

Daily county-level counts 1986-2013. Data location: https://www1.ncdc.noaa.gov/pub/data/swdi/reports/county/byFips/
Note: this data is not spatial and cannot be bounded by the ANF.For exploratory analysis purposes, this data will be explored across counties that the ANF is within. These counties are Liberty, Wakulla, Franklin, and Leon.

First get the Florida fips codes for Liberty, Wakulla, Franklin and Leon counties, then get the data.
```{r}
library(tidycensus)

FLfips <- fips_codes %>%
  filter(state == "FL") %>%
  filter(county %in% c("Liberty County", "Wakulla County", "Franklin County", "Leon County")) %>%
  pull(county_code)

fn <- paste0("https://www1.ncdc.noaa.gov/pub/data/swdi/reports/county/byFips/swdireport-12", FLfips, "-BETA.csv")

lightningdata.df <- data.frame()
for(i in 1:length(fn)){
  X <- read.csv(fn[i], na.strings = "NULL", header = TRUE, stringsAsFactors = FALSE)
  lightningdata.df <- rbind(lightningdata.df, X)
}

lightningdata.df <- lightningdata.df %>%
  mutate(SEQDAY = as.Date(SEQDAY),
         DAY = day(SEQDAY),
         MONTH = month(SEQDAY),
         YEAR = year(SEQDAY))
```

lightningdata.df gives lighting data for each individual county based on fips code. Combine total lightning strikes in ANF counties by date.
```{r}
countlightning.df <- lightningdata.df %>%
  group_by(SEQDAY) %>%
  summarise(FCOUNT_NLDN = sum(FCOUNT_NLDN)) %>%
  rename(ANFcountiesCount = FCOUNT_NLDN)
```

Explore relationship between high lightning count days and number of lightning sparked wildfires in the ANF

Import Forest Boundaries to explore lightning firedata in ANF
```{r}
library(tidyverse)
library(sf)

if(!"S_USA.NFSLandUnit" %in% list.files()){
  download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip",
                "S_USA.NFSLandUnit.zip")
unzip("S_USA.NFSLandUnit.zip")
}

NF_Bounds.sf <- st_read(dsn = "s_USA.NFSLandUnit.shp") %>%
  st_transform(crs = 3086)

anfbounds.sf <- NF_Bounds.sf %>%
  filter(NFSLANDU_2 == "Apalachicola National Forest")
```

Import fire data
```{r}
if(!"FL_Fires" %in% list.files()){
download.file("http://myweb.fsu.edu/jelsner/temp/data/FL_Fires.zip",
"FL_Fires.zip")
unzip("FL_Fires.zip")
}
FL_Fires.sf <- st_read(dsn = "FL_Fires") %>%
st_transform(crs = 3086)

anf_fires.sf <- st_join(FL_Fires.sf, anfbounds.sf, join = st_within) %>%
  filter(NFSLANDU_2 == 'Apalachicola National Forest')
```

Filtered for Lightning
```{r}
anf_LF.sf <- anf_fires.sf %>%
  filter(STAT_CAU_1 == "Lightning")
```

create data set containing fires and lightning counts. Match yaers for data sets (2005 - 2013)
```{r}
exploreLF <- anf_LF.sf %>%
  filter(FIRE_Y <= 2013)
exploreCount <- countlightning.df %>%
  filter(year(SEQDAY) >= 1992)


exploratory <- merge(x = anf_LF.sf, y = countlightning.df, by.x = c("DISCOVERY_"), by.y = c("SEQDAY"))
```

Explore if high lightning count days have fires.
```{r}

```




