---
title: "ExploratoryAnalysis_Committee"
author: "Zach"
date: "5/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
This file will explore ideas proposed by the committee during the prospectus defense on 5/25/2021


Import Libraries to be used
```{r}
library(lubridate)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(tidycensus)
library(dplyr)
library(readr)
library(sf)
library(tmap)
library(rgeos)
library(RColorBrewer)
```




### Checking Value of Q

It is expected that with time KBDI plots will become equal regardless of the initial value of Q


Copy and paste KBDI code

## Import the summary of the day data and add columns to the data frame

Assign no rainfall on days with missing values.

For Tallahassee Station TLH
```{r}
TLH.df <- read_csv(file = 'Data/TLH_Daily1940.csv') %>%
  rename(Date = DATE) %>%
  mutate(Year = year(Date), 
         month = month(Date, label = TRUE, abbr = TRUE),
         doy = yday(Date),
         MaxTemp = TMAX,
         MinTemp = TMIN,
         Rainfall24 = PRCP,
         Rainfall24 = replace_na(Rainfall24, 0),
         Rainfall24mm = Rainfall24 * 25.4)

#Fill in the missing temperature data
TLH.df$MaxTemp[TLH.df$Date == "2005-07-08"] <- 96
```


A new data frame is created to compare initial Q values. Q is representative of the starting value of KBDI. Here a column of several Q values is created to be used and compared among multiple plots. Daily KBDI is calculated for each initial Q value. This is testing the theory that with enough time elapsed, the initial value of Q does not matter.

rangeQ.df data frame is initialized at start of TLH.df data frame. (March 1940)
```{r}
Rainfall24 <- TLH.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0
CumR <- 0
NetR <- numeric()
for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH.df$NetR <- NetR
R <- 59.23

rangeQ.df <- TLH.df %>%
  select(Date, Rainfall24, MaxTemp, NetR)

for(i in 1:9){
  Q2 <- i-1
  Q <- Q2*100
  MaxTemp <- TLH.df$MaxTemp
  Ql <- numeric()
  for(i in 1:length(Rainfall24)){
    DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
    Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
    Q <- ifelse(Q < 0, 0, Q) 
    Ql <- c(Ql, Q)
  }
  rangeQ.df[, ncol(rangeQ.df)+1] <- Ql
  colnames(rangeQ.df) <- c("Date", "Rainfall24", "MaxTemp", "NetR", "Q0", "Q100", "Q200", "Q300", "Q400", "Q500", "Q600", "Q700", "Q800")
}
```

```{r}
head(rangeQ.df)
```


Plot Q values for first year of data period (1940)
```{r}
p1_1940 <- rangeQ.df %>% 
  filter(year(Date) == 1940) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(aes(y = Q0, color = "0")) +
    geom_line(aes(y = Q400, color = "400")) + 
    geom_line(aes(y = Q800, color = "800")) +
    scale_color_manual("Initial Q-values",
                       breaks = c("0", "400", "800"),
                       values = c("red", "blue", "black")) +
  ylab("KBDI") +
  theme(legend.position = "bottom")
  #theme(legend.position = c(0.8,0.3)

p2_1940 <- rangeQ.df %>% 
  filter(year(Date) == 1940) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(mapping = aes(y = Rainfall24)) +
  ylab("24Hr Net Rainfall")

p1_1940 + p2_1940 +
  plot_annotation(
    title = "KBDI values merge with time regardless of initial Q-values",
    subtitle = "Q-values intialized at 0, 400, and 800",
    caption = "Period of Record: March 1940 - December 1940") 
```


Filtering TLH data frame to start in 1991. This is one year prior to the fire data obtained from the Forest Service Research Data Archive
```{r}
TLH1991.df <- TLH.df %>%
  filter(year(Date) >= 1991)
```

Initialize different Q values starting in 1991 and create rangeQ1991.df
```{r}
Rainfall24 <- TLH1991.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0
CumR <- 0
NetR <- numeric()
for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH1991.df$NetR <- NetR
R <- 59.23

rangeQ1991.df <- TLH1991.df %>%
  select(Date, Rainfall24, MaxTemp, NetR)

for(i in 1:9){
  Q2 <- i-1
  Q <- Q2*100
  MaxTemp <- TLH1991.df$MaxTemp
  Ql <- numeric()
  for(i in 1:length(Rainfall24)){
    DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
    Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
    Q <- ifelse(Q < 0, 0, Q) 
    Ql <- c(Ql, Q)
  }
  rangeQ1991.df[, ncol(rangeQ1991.df)+1] <- Ql
  colnames(rangeQ1991.df) <- c("Date", "Rainfall24", "MaxTemp", "NetR", "Q0", "Q100", "Q200", "Q300", "Q400", "Q500", "Q600", "Q700", "Q800")
}
```

Plot 1991, year before fire data starts
```{r}
p1_1991 <- rangeQ1991.df %>% 
  filter(year(Date) == 1991) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(aes(y = Q0, color = "0")) +
    geom_line(aes(y = Q400, color = "400")) + 
    geom_line(aes(y = Q800, color = "800")) +
    scale_color_manual("Initial Q-values",
                       breaks = c("0", "400", "800"),
                       values = c("red", "blue", "black")) +
  ylab("KBDI") +
  theme(legend.position = "bottom")
  #theme(legend.position = c(0.8,0.3)

p2_1991 <- rangeQ1991.df %>% 
  filter(year(Date) == 1991) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(mapping = aes(y = Rainfall24)) +
  ylab("24Hr Net Rainfall")

p1_1991 + p2_1991 +
  plot_annotation(
    title = "KBDI values merge with time regardless of initial Q-values",
    subtitle = "Q-values intialized at 0, 400, and 800",
    caption = "Period of Record: 1991") 
```

focus plot on first month
```{r}
p1_1991 <- rangeQ1991.df %>% 
  filter(year(Date) == 1991) %>%
  filter(month(Date) == 1) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(aes(y = Q0, color = "0")) +
    geom_line(aes(y = Q400, color = "400")) + 
    geom_line(aes(y = Q800, color = "800")) +
    scale_color_manual("Initial Q-values",
                       breaks = c("0", "400", "800"),
                       values = c("red", "blue", "black")) +
  ylab("KBDI") +
  theme(legend.position = "bottom")
  #theme(legend.position = c(0.8,0.3)

p2_1991 <- rangeQ1991.df %>% 
  filter(year(Date) == 1991) %>%
  filter(month(Date) == 1) %>%
  ggplot(mapping = aes(x = Date)) +
    geom_line(mapping = aes(y = Rainfall24)) +
  ylab("24Hr Net Rainfall")

p1_1991 + p2_1991 +
  plot_annotation(
    title = "KBDI values merge with time regardless of initial Q-values",
    subtitle = "Q-values intialized at 0, 400, and 800",
    caption = "Period of Record: January 1991") 
```


##Explore lightning data and concerns mentioned by committee.

Get lightning data

Daily county-level counts 1986-2013. Data location: https://www1.ncdc.noaa.gov/pub/data/swdi/reports/county/byFips/
Note: this data is not spatial and cannot be bounded by the ANF.For exploratory analysis purposes, this data will be explored across counties that the ANF is within. These counties are Liberty, Wakulla, Franklin, and Leon.

First get the Florida fips codes for Liberty, Wakulla, Franklin and Leon counties, then get the data.
```{r}
FLfips <- fips_codes %>%
  filter(state == "FL") %>%
  filter(county %in% c("Liberty County", "Wakulla County", "Franklin County", "Leon County")) %>%
  pull(county_code)

fn <- paste0("https://www1.ncdc.noaa.gov/pub/data/swdi/reports/county/byFips/swdireport-12", FLfips, "-BETA.csv")

lightningdata.df <- data.frame()
for(i in 1:length(fn)){
  X <- read.csv(fn[i], na.strings = "NULL", header = TRUE, stringsAsFactors = FALSE)
  lightningdata.df <- rbind(lightningdata.df, X)
}

lightningdata.df <- lightningdata.df %>%
  mutate(SEQDAY = as.Date(SEQDAY),
         DAY = day(SEQDAY),
         MONTH = month(SEQDAY),
         YEAR = year(SEQDAY))
```

Attempt to read files from each downloaded csv because website import is not reliable and does not always open
```{r}
lightningdata1.df <- list.files(path = "C:/Users/zlaw9/OneDrive/GITHUB/KDBI code/FIPS_LightningData",
                  pattern = "*.csv", full.names = TRUE) %>%
  lapply(read_csv) %>%
  bind_rows

lightningdata1.df <- lightningdata1.df %>%
  mutate(SEQDAY = as.Date(SEQDAY),
         DAY = day(SEQDAY),
         MONTH = month(SEQDAY),
         YEAR = year(SEQDAY)) %>%
#missing data beyond 5/20/2013 for FIPS #12037. Remove last 5 rows from data set. This addresses the parsing failure and removes data that is NA.
  filter(SEQDAY <=  "2013-05-20")
```


lightningdata1.df gives lighting data for each individual county based on fips code. Combine total lightning strikes in ANF counties by date.

```{r}
countlightning.df <- lightningdata1.df %>%
  group_by(SEQDAY) %>%
  summarise(LightningCount = sum(FCOUNT_NLDN))
```

Explore the relationship between high lightning count days and number of lightning sparked wildfires in the ANF

County boundaries will be used instead of forest boundaries to keep bounds of fires and lightning consistent. Run county bounds code chunk.
Import Forest Boundaries to explore lightning fire data bounded by the ANF
```{r}
if(!"S_USA.NFSLandUnit" %in% list.files()){
  download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip",
                "S_USA.NFSLandUnit.zip")
unzip("S_USA.NFSLandUnit.zip")
}

NF_Bounds.sf <- st_read(dsn = "s_USA.NFSLandUnit.shp") %>%
  st_transform(crs = 3086)

anfbounds.sf <- NF_Bounds.sf %>%
  filter(NFSLANDU_2 == "Apalachicola National Forest")
```

Import County Boundaries of the forest. This will make the bounds consistent across both the fire data set and the lightning data set.
```{r}
ANFcountyBounds.sf <- st_read(dsn = "ANFCounties.shp") %>%
  st_transform(crs = 3086)
```



Import fire data
```{r}
if(!"FL_Fires" %in% list.files()){
download.file("http://myweb.fsu.edu/jelsner/temp/data/FL_Fires.zip",
"FL_Fires.zip")
unzip("FL_Fires.zip")
}
FL_Fires.sf <- st_read(dsn = "FL_Fires") %>%
st_transform(crs = 3086)

#filtered fires within ANF bounds
#anf_fires.sf <- st_join(FL_Fires.sf, anfbounds.sf, join = st_within) %>%
  #filter(NFSLANDU_2 == 'Apalachicola National Forest')

#filtered for county bounds of ANF. This is to match the bounds of the lightning dataset
anf_fires.sf <- st_join(FL_Fires.sf, ANFcountyBounds.sf, join = st_within) %>%
  filter(CNTY_FIPS != "NA") %>%
  select(FOD_I, FIRE_N, FIRE_Y, DISCOVERY_, STAT_CAU_1, FIRE_SIZE, FIRE_SIZE_, LATIT, LONGI, NAME, FIPS, geometry)

anf_LF.sf <- anf_fires.sf %>%
  filter(STAT_CAU_1 == "Lightning")
```

Archived fire data has been updated through 2018. Use this code chunk for most up to data fire file.
Website link with data and metadata
https://www.fs.usda.gov/rds/archive/Catalog/RDS-2013-0009.5 

Fire set downloaded and pulled from file rather than directly from the website. Had errors when trying to import it directly from the website because the file format could not be read. Downloaded and converted to more user friendly shapefiles in ESRI.

New Data Set has change categories for cause. Lightning is not listed. Is lightning the equivalent of fires filtered as natural? Explore this at the bottom of the rmd.
```{r}
updatedFires.sf <- st_read(dsn = "C:/Users/zlaw9/OneDrive/GITHUB/KDBI code/updatedFires",
                           layer = "FLFiresUpdated") %>%
  st_transform(crs = 3086)

#filtered for county bounds of ANF. This is to match the bounds of the lightning dataset
anf_fires.sf <- st_join(updatedFires.sf, ANFcountyBounds.sf, join = st_within) %>%
  filter(CNTY_FIPS != "NA") %>%
  select(FOD_I, FIRE_N, FIRE_Y, DISCOVERY_, NWCG_CAUSE_CLASSIFICATION, NWCG_GENERAL_CAUSE, FIRE_SIZE, FIRE_SIZE_, LATIT, LONGI, NAME, FIPS, geometry)

anf_LF.sf <- anf_fires.sf %>%
  filter(STAT_CAU_1 == "Lightning")
```



View fire points that have been bounded by four counties (Liberty, Leon, Wakulla, and Franklin)
```{r}
tmap_mode("view")

tm_shape(ANFcountyBounds.sf) +
  tm_borders()

tm_shape(anf_LF.sf) +
  tm_dots(col = "orange")
```

create data set containing fires and lightning counts. Match the years across the data sets (1992 - 2013)
Merge1.df is a new data frame that contains the number of lightning strikes and fires occurring in the Apalachicola National Forest broken down by day from 1992 - 2013.
```{r}
#exploreLF <- anf_LF.sf %>%
  #filter(FIRE_Y <= 2013)

countfires.df <- anf_LF.sf %>%
#countfires.df <- exploreLF %>%
  count(DISCOVERY_) %>%
  rename(FireCount = n, Date = DISCOVERY_)

countlightning.df <- countlightning.df %>%
  filter(year(SEQDAY) >= 1992)
  #rename(LightningCount = FCOUNT_NLDN)

merge1.df <- left_join(x = countlightning.df, y = countfires.df, by = c("SEQDAY" = "Date")) %>%
  rename(Date = SEQDAY) %>%
  select(Date, LightningCount, FireCount)
```

Recalculate QLm for intital Q value of 269. Create new data frame (TLH1.df) to be merged with lightningcount/firecount dataframe (merge1)
```{r}
TLH1.df <- TLH.df %>%
  filter(year(Date) >= 1992) %>%
  filter(Date <= "2013-05-25")
  
Rainfall24 <- TLH1.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0

CumR <- 0
NetR <- numeric()

for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH1.df$NetR <- NetR

Q <- 269  
R <- 59.23 # average annual rainfall for TLH in inches

MaxTemp <- TLH1.df$MaxTemp

Ql <- numeric()
DeltaQl <- numeric()
for(i in 1:length(Rainfall24)){
  DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
  Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
  Q <- ifelse(Q < 0, 0, Q) 
  Ql <- c(Ql, Q)
  DeltaQl <- c(DeltaQl, DeltaQ)
}

TLH1.df$Ql <- Ql
TLH1.df$Qlm <- Ql * .254  # tenth of an inch to mm
TLH1.df$DeltaQl <- DeltaQl
TLH1.df$DroughtIndex <- floor(Ql/100)
```



Merge data frames to have lightning strike count, fire count and Qlm all in one dataframe
```{r}
merge2.df <- left_join(x = merge1.df, y = TLH1.df, by = c("Date")) %>%
  select(Date, LightningCount, FireCount, Ql)

merge2.df$FireBool <- !is.na(merge2.df$FireCount)
```

General linear regression - Study how lightning influences the probability of fire occurrence 
Predict the occurrence of a fire for each day in the data set based on lightning count.
Note: This plot includes days with zero lightning strikes and concludes that on days with zero lightning strikes there is a 3% chance of a lightning ignited wildfire occurring. A lightning fire cannot occur without lighting. Should be zero percent.
```{r}
#general linear regression
glmLightning <- glm(formula = FireBool ~ LightningCount, family = binomial, data = merge2.df)

#probabilities based on glm
pred_glmLightning <- predict(object = glmLightning,
                             type = "response",
                             se.fit = TRUE)
#creating confidence interval
lowerLightning <- pred_glmLightning$fit - (1.96*pred_glmLightning$se.fit)
upperLightning <- pred_glmLightning$fit + (1.96*pred_glmLightning$se.fit)

ggplot(mapping = aes(x = merge2.df$LightningCount, y = pred_glmLightning$fit)) +
  geom_ribbon(aes(ymin = lowerLightning, ymax = upperLightning), fill = "grey") +
  geom_line(color = "blue") +
  ylab("Predicited Probablility") +
  xlab("Number of Lightning Strikes") +
  labs(
    title = "Predicted Probability of a Fire Occurance Based on the Number of Lightning Strikes in a Day",
    subtitle = "With a 95% Confidence Interval",
    caption = "Period of Record: 1-1-1992 to 5-20-2013")
```


Filter data for only dates in the fire season before running glm.

When filtering to only the fire season, there is more uncertainty, and the model is closer to linear.
```{r}
fireSeasondates.df <- merge2.df %>%
  filter(month(Date) == 05 | month(Date) == 06 | month(Date) == 07)

#general linear regression
glmLightningFS <- glm(formula = FireBool ~ LightningCount, family = binomial, data = fireSeasondates.df)

#probabilities based on glm
pred_glmLightningFS <- predict(object = glmLightningFS,
                             type = "response",
                             se.fit = TRUE)
#creating confidence interval
lowerLightningFS <- pred_glmLightningFS$fit - (1.96*pred_glmLightningFS$se.fit)
upperLightningFS <- pred_glmLightningFS$fit + (1.96*pred_glmLightningFS$se.fit)

ggplot(mapping = aes(x = fireSeasondates.df$LightningCount, y = pred_glmLightningFS$fit)) +
  geom_ribbon(aes(ymin = lowerLightningFS, ymax = upperLightningFS), fill = "grey") +
  geom_line(color = "blue") +
  ylab("Predicited Probablility") +
  xlab("Number of Lightning Strikes") +
  labs(
    title = "Predicted Probability of Fire Occurance Based on the Number of Lightning Strikes in a Day\nFiltered By Fire Season Months",
    subtitle = "With 95% Confidence Interval",
    caption = "Fire Season Months (May - July) 1992 - 2013")
```



General linear regression - Study how KBDI impacts lightning wildfire occurrence.
```{r}
glmQl <- glm(formula = FireBool ~ Ql, family = binomial, data = merge2.df)

pred_glmQl <- predict(object = glmQl,
                      type = "response",
                      se.fit = TRUE)

lowerQl <- pred_glmQl$fit - (1.96*pred_glmQl$se.fit)
upperQl <- pred_glmQl$fit + (1.96*pred_glmQl$se.fit)

ggplot(mapping = aes(x = merge2.df$Ql, y = pred_glmQl$fit)) +
  geom_ribbon(aes(ymin = lowerQl, ymax = upperQl), fill = "grey") +
  geom_line(color = "blue") +
  ylab("Predicited Probablility") +
  xlab("Daily Ql") +
  labs(
    title = "Predicted Probability of Fire Occurance Based on Daily KBDI Values",
    subtitle = "With 95% Confidence Interval",
    caption = "Period of Record: 1-1-1992 to 5-20-2012")
```

KBDI impacts on lightning wildfire occurrence during the fire season. KBDI has a much stronger impact during the fire season months.
```{r}
glmQlFS <- glm(formula = FireBool ~ Ql, family = binomial, data = fireSeasondates.df)

pred_glmQlFS <- predict(object = glmQlFS,
                      type = "response",
                      se.fit = TRUE)

lowerQlFS <- pred_glmQlFS$fit - (1.96*pred_glmQlFS$se.fit)
upperQlFS <- pred_glmQlFS$fit + (1.96*pred_glmQlFS$se.fit)

ggplot(mapping = aes(x = fireSeasondates.df$Ql, y = pred_glmQlFS$fit)) +
  geom_ribbon(aes(ymin = lowerQlFS, ymax = upperQlFS), fill = "grey") +
  geom_line(color = "blue") +
  ylab("Predicited Probablility") +
  xlab("Daily Ql") +
  labs(
    title = "Predicted Probability of Fire Occurance Based on Daily KBDI Values\nFiltered By Fire Season Months",
    subtitle = "With 95% Confidence Interval",
    caption = "Fire Season Months (May - July) 1992 - 2013")
```



Run general linear regression. The number of lightning strikes and the saturation of the soil are both statistically significant.
```{r}
GLM <- glm(formula = FireBool ~ LightningCount + Ql, family = binomial, data = merge2.df)

pred_GLM <- predict(object = GLM,
                    type = "response",
                    se.fit = TRUE)

ggplot(mapping = aes(x = merge2.df$Ql, y = merge2.df$LightningCount, color = pred_GLM$fit)) +
  geom_point() +
  scale_color_gradient(low = "orange", high = "red4") +
  xlab("KBDI") +
  ylab("Number of Lightning Strikes") +
  labs(color = "Probability of Fire") +
  labs(title = "Predicted Probability of Fire Occurance\nBased on Daily KBDI Values and Daily Lightning Strikes",
       caption = "Period of Record: 1-1-1992 to 5-20-2012")
  

#GLM
#summary(GLM)
#exp(GLM$coefficients)
```

Filter for fire season
```{r}
GLMFS <- glm(formula = FireBool ~ LightningCount + Ql, family = binomial, data = fireSeasondates.df)

pred_GLMFS <- predict(object = GLMFS,
                    type = "response",
                    se.fit = TRUE)

ggplot(mapping = aes(x = fireSeasondates.df$Ql, y = fireSeasondates.df$LightningCount, color = pred_GLMFS$fit)) +
  geom_point() +
  scale_color_gradient(low = "orange", high = "red4") +
  xlab("KBDI") +
  ylab("Number of Lightning Strikes") +
  labs(color = "Probability of Fire") +
  labs(title = "Predicted Probability of Fire Occurance\nBased on Daily KBDI Values and Daily Lightning Strikes",
       caption = "Fire Season Months (May - July) 1992 - 2013")
```


Explore if the occurrence of a large fire creates a time lag in the occurrence of another fire.
Explore relationship between the number of fires and fire size.

```{r}
seasonFires <- anf_LF.sf %>%
  filter(month(DISCOVERY_) == 05 | month(DISCOVERY_) == 06 | month(DISCOVERY_) == 07) %>%
  count(FIRE_Y) %>%
  rename(Year = FIRE_Y, nFIRES = n) %>%
  st_set_geometry(NULL)

LargestFires <- anf_LF.sf %>%
  filter(month(DISCOVERY_) == 05 | month(DISCOVERY_) == 06 | month(DISCOVERY_) == 07) %>%
  group_by(FIRE_Y) %>%
  summarise(LargestFire = max(FIRE_SIZE)) %>%
  rename(Year = FIRE_Y) %>%
  st_set_geometry(NULL)

seasonSummary <- merge(x = seasonFires, y = LargestFires, by = c("Year"))

head(seasonSummary)

#cor(seasonSummary$nFIRES, seasonSummary$LargestFire)
```



###Explore updated fire data set compared to the old data set

Filters have changed between the new data set, and fires are not defined by cause = lightning. Natural is introduced as a new classification. Are lightning fires the equivalent of natural fires?

Per meta data (copy and pasted):
NWCG_CAUSE_CLASSIFICATION = Broad classification of the reason the fire occurred (Human, Natural, Missing data/not specified/undetermined).

NWCG_GENERAL_CAUSE = Event or circumstance that started a fire or set the stage for its occurrence (Arson/incendiarism, Debris and open burning, Equipment and vehicle use, Firearms and explosives use, Fireworks, Misuse of fire by a minor, Natural, Power generation/transmission/distribution, Railroad operations and maintenance, Recreation and ceremony, Smoking, Other causes, Missing data/not specified/undetermined).


Create est simple features for old data
```{r}
if(!"FL_Fires" %in% list.files()){
download.file("http://myweb.fsu.edu/jelsner/temp/data/FL_Fires.zip",
"FL_Fires.zip")
unzip("FL_Fires.zip")
}
FL_FiresTest1.sf <- st_read(dsn = "FL_Fires") %>%
st_transform(crs = 3086)

#filtered fires within ANF bounds
#anf_fires.sf <- st_join(FL_Fires.sf, anfbounds.sf, join = st_within) %>%
  #filter(NFSLANDU_2 == 'Apalachicola National Forest')

#filtered for county bounds of ANF. This is to match the bounds of the lightning dataset
anf_firesTest1.sf <- st_join(FL_FiresTest1.sf, ANFcountyBounds.sf, join = st_within) %>%
  filter(CNTY_FIPS != "NA") %>%
  select(FOD_I, FIRE_N, FIRE_Y, DISCOVERY_, STAT_CAU_1, FIRE_SIZE, FIRE_SIZE_, LATIT, LONGI, NAME, FIPS, geometry)

anf_LFTest1.sf <- anf_firesTest1.sf %>%
  filter(STAT_CAU_1 == "Lightning") %>%
  #drop geometry so data sets can merge
  st_drop_geometry()
```

Create test simple features for new data
```{r}
updatedFiresTest2.sf <- st_read(dsn = "C:/Users/zlaw9/OneDrive/GITHUB/KDBI code/updatedFires",
                           layer = "FLFiresUpdated") %>%
  st_transform(crs = 3086)

#filtered for county bounds of ANF. This is to match the bounds of the lightning dataset
anf_firesTest2.sf <- st_join(updatedFiresTest2.sf, ANFcountyBounds.sf, join = st_within) %>%
  filter(CNTY_FIPS != "NA") %>%
  select(FOD_ID, FIRE_NAME, FIRE_YEAR, DISCOVERY_, NWCG_CAUSE, NWCG_GENER, FIRE_SIZE, FIRE_SIZE_, LATITUDE, LONGITUDE, COUNTY, FIPS_CODE, geometry)

anf_LFTest2.sf <- anf_firesTest2.sf %>%
  filter(NWCG_GENER == "Natural")
```

Join two data sets together to do bolean comparison. Boolean column is all TRUE. Therefore we can conclude that fires filtered as natural are the equivalent of fires filtered as Lightning in the 2015 data set.
```{r}
TestFires.sf <- left_join(x = anf_LFTest2.sf, y = anf_LFTest1.sf, by = c("FOD_ID" = "FOD_I")) %>%
  select(FOD_ID, FIRE_NAME, FIRE_N, FIRE_YEAR, FIRE_Y, NWCG_GENER, STAT_CAU_1) %>%
  rename(NAME2018 = FIRE_NAME, NAME2015 = FIRE_N, YEAR2018 = FIRE_YEAR, YEAR2015 = FIRE_Y, CAUSE2018 = NWCG_GENER, CAUSE2015 = STAT_CAU_1)

#create bool argeument to test if Lightining Fires are equivalent to Natural Fires
TestFires.sf$BOOL <- TestFires.sf$CAUSE2018 == "Natural" & TestFires.sf$CAUSE2015 == "Lightning"
```

Count number of new data points added in the 2018 set. There are 30 new data points.
```{r}
sum(is.na(TestFires.sf$YEAR2015))
```


###Still Working on this
Further exploring fire and drought data before starting larger regression models:
Plot data and explore trends.

```{r}
Rainfall24 <- TLH.df$Rainfall24
PR <- dplyr::lag(Rainfall24)
PR[1] <- 0

CumR <- 0
NetR <- numeric()

for(i in 1:length(Rainfall24)) {
  R24 <- Rainfall24[i]
  if (R24 == 0) {
    NetR[i] <- 0
    CumR <- 0
  } 
  else if(R24 > 0 & R24 <= .2) {
      CumR <- CumR + R24
      if (PR[i] > .2 | CumR > .2) NetR[i] <- R24
      else if (CumR > .2) NetR[i] <- CumR - .2
      else NetR[i] <- 0
    }
  else if (R24 > .2) {
      if (CumR <= .2) {
      NetR[i] <- CumR + R24 - .2
      CumR <- CumR + R24
      }
      else {
      NetR[i] <- R24
      CumR <- CumR + R24
      }
  }
}

TLH.df$NetR <- NetR

Q <- 269  
R <- 59.23 # average annual rainfall for TLH in inches

MaxTemp <- TLH.df$MaxTemp

Ql <- numeric()
DeltaQl <- numeric()
for(i in 1:length(Rainfall24)){
  DeltaQ <- (800 - Q) * (.968 * exp(.0486 * MaxTemp[i]) - 8.3) /(1 + 10.88 * exp(-.0441 * R)) * .001 
  Q <- ifelse(NetR[i] == 0,  Q + DeltaQ,  (Q + DeltaQ) - NetR[i] * 100)
  Q <- ifelse(Q < 0, 0, Q) 
  Ql <- c(Ql, Q)
  DeltaQl <- c(DeltaQl, DeltaQ)
}

TLH.df$Ql <- Ql
TLH.df$Qlm <- Ql * .254  # tenth of an inch to mm
TLH.df$DeltaQl <- DeltaQl
TLH.df$DroughtIndex <- floor(Ql/100)

TLH.df <- TLH.df %>%
 dplyr::filter(Year >= 1946 & Year <= 2019) # Only full years
```

```{r}
droughtFire <- merge(x = anf_LF.sf, y = TLH.df, by.x = c("DISCOVERY_"), by.y = c("Date"))

droughtFire <- droughtFire %>%
  select("Date", "Qlm", "FIRE_SIZE")
```








