---
title: "Generative model"
output: html_document
editor_options: 
  chunk_output_type: console
---

Create a generative model for natural wildfire risk.

Use McElreath's {rethinking} package. See https://github.com/rmcelreath/rethinking

```{r}
devtools::install_github("rmcelreath/rethinking")
```

Almost any ordinary generalized linear model can be specified with `quap()`. To use quadratic approximation:
```{r}
library(rethinking)

f <- alist(
    y ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10),
    sigma ~ dexp(1)
)

fit <- quap( 
    f, 
    data = list(y = c(-3, 2)), 
    start = list(mu = 0, sigma = 1)
)

precis(fit)
```

This tells us that with two data values y = -3 and 2 the best estimate for mu is -.49 and the best estimate for sigma is 1.81, given the model.

The object `fit` holds the result. For a summary of marginal posterior distributions, use `summary(fit)` or `precis(fit)`.
```{r}
summary(fit)
```

We simulate posterior values with the `sim()` function.
```{r}
sim(fit, n = 20)
```

Each row is a posterior sample and each column is an element of the data vector (2-D) since there are only two data values.

Poisson counts.
```{r}
f2 <- alist(
  y ~ dpois(lambda),
  lambda ~ dgamma(shape = 2, scale = 10)
)

fit2 <- quap( 
    f2, 
    data = list(y = c(0, 1, 2, 3)), 
    start = list(lambda = 1)
)

precis(fit2)
```

The data are organized in the file `ANF_Fires.Rmd`. 
```{r}
library(tidyverse)

SeasonalData.df <- read_csv(file = "SeasonalForecastModelData.csv") %>%
  mutate(QlmLastDay = QlmLastDay / 10) # change units from mm to cm

W <- SeasonalData.df$nFires

f3 <- alist(
  y ~ dpois(lambda),
  lambda ~ dgamma(shape = 100, scale = 40)
)

fit3 <- quap( 
    f3, 
    data = list(y = W), 
    start = list(lambda = 1)
)

precis(fit3)

extract.samples(fit3, n = 10)
sim(fit3, n = 2)
```

Hamiltonian Monte Carlo with `ulam()`.

It is named Stanisław Ulam, who was one of the parents of the Monte Carlo method and is the namesake of the Stan project as well. It is pronounced something like [OO-lahm], not like [YOU-lamm].

It takes the same kind of input as `quap()`.
```{r}
fit_stan <- ulam(f3,
                 data = list(y = W) )
```