---
title: "Gridded daily weather data"
output: html_document
editor_options: 
  chunk_output_type: console
---

Can we do better than the KBDI computed at TLH using daily gridMET data (burn index, 1000 hr moist) over the ANF when predicting seasonal lightning-sparked wildfires?

```{r}
# remotes::install_github("mikejohnson51/AOI")
# remotes::install_github("mikejohnson51/climateR")

library(tidyverse)
library(lubridate)
library(rgdal)
library(sf)
library(raster)
library(AOI)
library(climateR)
select <- dplyr::select  # make sure that the select you want is from {dplyr} (not the one in {raster})
```

See `ClimateR.Rmd` and http://www.climatologylab.org/gridmet.html and https://youtu.be/OhxAISoi2zE

What variables are available relevant to wildfires.
```{r}
param_meta$gridmet
```

Get the boundary file.
```{r}
if(!"S_USA.NFSLandUnit" %in% list.files()){
  download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip",
                "S_USA.NFSLandUnit.zip")
unzip("S_USA.NFSLandUnit.zip",
      files = "S_USA.NFSLandUnit")
}

ANF_Boundary.sf <- st_read(dsn = "S_USA.NFSLandUnit") %>%
  filter(NFSLANDU_2 == "Apalachicola National Forest")

st_area(ANF_Boundary.sf) # 2,564 sq. km (634,000 acre)
```

Get daily variables as `RasterStack`s from 1992-2015. About 12 seconds for 4 variables.
```{r}
t0 <- Sys.time()
data <- getGridMET(ANF_Boundary.sf, 
                   param = c("tmax", "prcp", "fmoist_1000", "burn_index"), 
                   startDate = "1992-01-01",
                   endDate = "2015-12-31")
Sys.time() - t0
class(data$gridmet_fmoist_1000)
nlayers(data$gridmet_fmoist_1000) # should equal number of days between start & end dates
```

Check to see if we have the correct region.
```{r}
plot(mean(data$gridmet_burn_index))
plot(ANF_Boundary.sf$geometry, add = TRUE)
```

Compute regional averages by date.
```{r}
regionalAvgFmoist <- cellStats(data$gridmet_fmoist_1000, stat = 'mean')
regionalAvgBurnIndex <- cellStats(data$gridmet_burn_index, stat = 'mean')
dailyDates <- seq(from = as.Date("1992/1/1"), 
                   to = as.Date("2015/12/31"), 
                   by = "day")
regional.df <- data.frame(dailyDates, 
                          regionalAvgFmoist,
                          regionalAvgBurnIndex) %>%
  mutate(Month = month(dailyDates),
         Year = year(dailyDates))
```

Get the KDBI data. The data are organized and merged in the file `ANF_Fires.Rmd`. 
```{r}
SeasonalData.df <- read_csv(file = "SeasonalForecastModelData.csv") %>%
  mutate(QlmLastDay = QlmLastDay / 10) # change units from mm to cm
```

Attach the regional data.
```{r}
df <- regional.df %>%
  group_by(Month, Year) %>%
  summarize(monthlyAvgFmoist = mean(regionalAvgFmoist),
            monthlyAvgBurnIndex = mean(regionalAvgBurnIndex),
            lastDayAvgFmoist = last(regionalAvgFmoist),
            lastDayAvgBurnIndex = last(regionalAvgBurnIndex)) %>%
  filter(Month == 4)

SeasonalData.df$monthlyAvgFmoist = df$monthlyAvgFmoist
SeasonalData.df$monthlyAvgBurnIndex = df$monthlyAvgBurnIndex
SeasonalData.df$lastDayAvgFmoist = df$lastDayAvgFmoist
SeasonalData.df$lastDayAvgBurnIndex = df$lastDayAvgBurnIndex
```

```{r}
cor(SeasonalData.df$nFires, SeasonalData.df$QlmLastDay)
# 0.6315948
cor(SeasonalData.df$nFires, SeasonalData.df$lastDayAvgFmoist)
#-0.6225665
cor(SeasonalData.df$nFires, SeasonalData.df$lastDayAvgBurnIndex)
# 0.2888821

cor(SeasonalData.df$nFires, SeasonalData.df$AvgQlm)
# 0.5821398
cor(SeasonalData.df$nFires, SeasonalData.df$monthlyAvgFmoist)
# -0.5238168
cor(SeasonalData.df$nFires, SeasonalData.df$monthlyAvgBurnIndex)
# 0.6597644

cor(SeasonalData.df$totalAcresBurned, SeasonalData.df$QlmLastDay)
cor(SeasonalData.df$totalAcresBurned, SeasonalData.df$lastDayAvgFmoist)
cor(SeasonalData.df$totalAcresBurned, SeasonalData.df$lastDayAvgBurnIndex)
cor(SeasonalData.df$totalAcresBurned, SeasonalData.df$AvgQlm)
cor(SeasonalData.df$totalAcresBurned, SeasonalData.df$monthlyAvgFmoist)
cor(SeasonalData.df$totalAcresBurned, SeasonalData.df$monthlyAvgBurnIndex)

# -0.4611462
# 0.3721229
# 0.5695327
# -0.5783408
# 0.6733335

library(MASS)

model5 <- glm.nb(nFires ~ monthlyAvgFmoist,
                 data = SeasonalData.df)
summary(model5)
```

KBDI computed from data at TLH is highly correlated with area-averaged Fmoist and burn index.

Get a the pentad (5-day) PDSI from 1980 to 2019.
```{r}
data <- getGridMET(ANF_Boundary.sf, 
                   param = "palmer", 
                   startDate = "1979-01-01",
                   endDate = "2020-01-02")
class(data$gridmet_palmer)

# Check that we get the correct region
Y <- mean(data$gridmet_palmer)
plot(Y)
plot(ANF_Boundary.sf$geometry, add = TRUE)

regionalAvgPDSI <- cellStats(data$gridmet_palmer, stat = 'mean')
pentadDates <- seq(from = as.Date("1980/1/3"), 
                   to = as.Date("2019/12/28"), 
                   by = "5 day")
PDSI.df <- data.frame(pentadDates, regionalAvgPDSI) %>%
  mutate(Month = month(pentadDates),
         Year = year(pentadDates))
```

