---
title: "Seasonal prediction of lightning-sparked wildfires in the Apalachicola National Forest (ANF)"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Get the required packages
```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(MASS)
library(brms)
library(tidybayes)
library(modelr)
```

## Get the data

The data are organized and merged in the file `ANF_Fires.Rmd`. 
```{r}
SeasonalData.df <- read_csv(file = "SeasonalForecastModelData.csv") %>%
  mutate(QlmLastDay = QlmLastDay / 10) # change units from mm to cm
```

Negative binomial regression.
```{r}
var(SeasonalData.df$nFires)/mean(SeasonalData.df$nFires) #14.2

model0 <- glm.nb(nFires ~ 1,
                 data = SeasonalData.df)
summary(model0)

model1 <- glm.nb(nFires ~ QlmLastDay + AprilSOI,
                 data = SeasonalData.df)
summary(model1)

model2 <- glm.nb(nFires ~ QlmLastDay,
                 data = SeasonalData.df)
summary(model2)

ggplot(data = SeasonalData.df, 
       mapping = aes(x = QlmLastDay, y = nFires)) +
  geom_point() + 
  geom_line(aes(y = fitted(model2))) +
  ylab("Number of Fires (May-July)") + xlab("Drought Index (April)") +
  theme_minimal()
```

Use the {brms} and {tidybayes} packages to get posterior distributions.

Also hot-days.Rmd, PredictTorFreqByEF.Rmd

When using brms, please cite one or more of the following publications:

Bürkner P. C. (2017). brms: An R Package for Bayesian Multilevel Models using Stan. Journal of Statistical Software. 80(1), 1-28. doi.org/10.18637/jss.v080.i01
Bürkner P. C. (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. The R Journal. 10(1), 395-411. doi.org/10.32614/RJ-2018-017
As brms is a high-level interface to Stan, please additionally cite Stan:

Carpenter B., Gelman A., Hoffman M. D., Lee D., Goodrich B., Betancourt M., Brubaker M., Guo J., Li P., and Riddell A. (2017). Stan: A probabilistic programming language. Journal of Statistical Software. 76(1). 10.18637/jss.v076.i01

```{r}
fit1 <- brm(nFires ~ QlmLastDay,
            family = "negbinomial",
            data = SeasonalData.df)

prior_summary(fit1)

plot(fit1, pars = "QlmLastDay")
```

For visualizing uncertain see https://www.youtube.com/watch?v=wbzfqh_3LyM&list=PLCrWEzJgSUqzI3goQEAKkDsHg72inmqbe&index=20&t=0s

```{r}
out <- SeasonalData.df %>%
  tidyr::expand(QlmLastDay = seq_range(QlmLastDay, n = 25)) %>% #GRID
  tidybayes::add_fitted_draws(fit1) %>% #CONDITION
  tidybayes::median_qi(.value, .width = c(.66, .95)) # MUNGE

ggplot(data = out,
       mapping = aes(x = QlmLastDay,
                     y = .value,
                     ymin = .lower,
                     ymax = .upper)) +
  geom_lineribbon()

SeasonalData.df %>%
  tidyr::expand(QlmLastDay = seq_range(QlmLastDay, n = 25)) %>% #GRID
  tidybayes::add_fitted_draws(fit1) %>% #CONDITION
  ggplot(mapping = aes(x = QlmLastDay,
                     y = .value)) +
  tidybayes::stat_lineribbon(.width = c(.66, .95))

                     
```
