---
title: "WeatherSTEM data across Leon County"
output: html_document
editor_options: 
  chunk_output_type: console
---

Data location: https://leon.weatherstem.com/

Choose a active WeatherSTEM site.
Get raw data by clicking on <raw data> under current conditions.

One station: https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/apalacheepark/davis.json

Get the data into R.
```{r}
library(tidyverse)

loc <- "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/apalacheepark/davis.json"
json_data <- fromJSON(file = loc)
```

Create a data frame
```{r}
ID <- json_data$davis_current_observation$DID
Name <- json_data$davis_current_observation$station_name
Latitude <- as.numeric(json_data$latitude)
Longitude <- as.numeric(json_data$longitude)
Day_Low_Time <- json_data$davis_current_observation$temp_day_low_time
Day_Low_Temp <- as.numeric(json_data$davis_current_observation$temp_day_low_f)

df <- data.frame(ID, Name, Latitude, Longitude, Day_Low_Temp, Day_Low_Time)
```

Create a station list.
```{r}
loc <- c("https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/apalacheepark/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/bucklake/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/canopyoaks/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/chaires/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/clc/davis.json",
#         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/chiles/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/conley/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/deerlake/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/desototrail/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/famudrs/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsus/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fortbraden/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsuesc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsursp/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsurez/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/gilchrist/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/hartsfield/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/hawksrise/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sullivan/davis.json", 
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/killearnlakes/davis.json",
          "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/leon/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/lincoln/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/maclay/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/netallahassee/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/oakridge/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/rickards/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/ruediger/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sail/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sealey/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/success/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/swiftcreek/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/tcc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/tallahasseenurseries/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/theedison/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/woodville/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/wtmoore/davis.json"
)
```

Loop over the station list importanting the data and making a data frame.
```{r}
ID <- NULL
Name <- NULL
Latitude <- NULL
Longitude <- NULL
Date <- NULL
Sunrise <- NULL
Day_Low_Time <- NULL
Day_Low_Temp <- NULL

for(i in 1:length(loc)){
  json_data <- fromJSON(file = loc[i])
  
    ID[i] <- json_data$davis_current_observation$DID
    Name[i] <- json_data$davis_current_observation$station_name
    Latitude[i] <- as.numeric(json_data$latitude)
    Longitude[i] <- as.numeric(json_data$longitude)
    Date[i] <- substring(json_data$observation_time_rfc822, 6, 16)
    Sunrise[i] <- json_data$davis_current_observation$sunrise
    Day_Low_Time[i] <- json_data$davis_current_observation$temp_day_low_time
    Day_Low_Temp[i] <- as.numeric(json_data$davis_current_observation$temp_day_low_f)
}

( df <- data.frame(ID, Name, Latitude, Longitude, Date, Sunrise, Day_Low_Time, Day_Low_Temp) %>%
    mutate(Date = as.Date(Date, format = "%d %B %Y")) %>%
    filter(Name != "DeSoto Trail Elementary WeatherSTEM") %>%
    filter(Date == as.Date("2020-12-11")) )
```

Convert the data frame to a simple feature data frame and make an interactive map.
```{r}
library(sf)

projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
sfdf <- st_as_sf(df, 
                 coords = c("Longitude", "Latitude"),
                 crs = projcrs)

library(tmap)

tmap_mode("view")

tm_shape(sfdf) +
  tm_dots() +
  tm_text(text = "Name")

tm_shape(sfdf) +
  tm_text(text = "Day_Low_Temp")
```

Get locations for interpolation.
```{r}
grid.sfc <- st_make_grid(sfdf,
                         n = c(20, 40),
                         what = "polygons")

tm_shape(grid.sfc) +
  tm_borders() +
tm_shape(sfdf) +
  tm_dots(col = "red")
```

Interpolate temperatures to a grid
```{r}
library(gstat)

interpolation <- idw(formula = Day_Low_Temp ~ 1,
                     locations = sfdf,
                     newdata = grid.sfc)

tm_shape(interpolation) +
  tm_fill(col = "var1.pred", alpha = .5)
```

Use projection first. Here we use Florida GDL Albers (EPSG:3087). This eliminates the warnings.
```{r}
sfdf <- st_transform(sfdf, 
                     crs = 3087)
grid.sfc <- st_make_grid(sfdf,
                         n = c(20, 40),
                         what = "polygons")
interpolation <- idw(formula = Day_Low_Temp ~ 1,
                     locations = sfdf,
                     newdata = grid.sfc)

tm_shape(interpolation) +
  tm_fill(col = "var1.pred", alpha = .5)
```
         