---
title: "WeatherSTEM data across Leon County"
output: html_document
editor_options: 
  chunk_output_type: console
---

Data location: https://leon.weatherstem.com/

Select stations and instruments then download.
```{r}
library(tidyverse)
df <- read.csv("Data/weatherSTEMdata2020-12-26.csv")
```

Organize the data. Start with the variable of interest. Here temperature.
```{r}
df2 <- df %>%
  select(contains("Thermometer")) %>%
  select(-c(contains("Chaires"), 
            contains("DeSoto"), 
            contains("Rickards"), 
            contains("Deerlake"),
            contains("Hawks"),
            contains("FSU.Weather"),
            contains("Montford"),
            contains("Golden.Eagle"),
            contains("Ballard"),
            contains("Florida.A.M."),
            contains("FSWN"),
            contains("Holy.Comforter")))
```

Next get the times as a vector.
```{r}
library(lubridate)

times <- ymd_hms(df$Timestamp)
```

Next create a station list in order of the data frame.
```{r}
names(df2)

loc <- c("https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/canopyoaks/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/leon/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/gilchrist/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sail/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/chiles/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/woodville/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fortbraden/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/lincoln/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sullivan/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/tallahasseenurseries/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/maclay/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/ruediger/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/conley/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/hartsfield/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/oakridge/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sealey/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/killearnlakes/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/wtmoore/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/bucklake/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/swiftcreek/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/famudrs/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsurez/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsursp/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/clc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsus/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/success/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/theedison/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/tcc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsuesc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/apalacheepark/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/netallahassee/davis.json"
)
```

Loop over the station list importing the latitude and longitude and making a data frame.
```{r}
Latitude <- rep(NA, length(loc))
Longitude <- rep(NA, length(loc))
Name <- rep(NA, length(loc))

for(i in 1:length(loc)){
  json_data <- fromJSON(file = loc[i])
    Name[i] <- json_data$davis_current_observation$station_name
    Latitude[i] <- as.numeric(json_data$latitude)
    Longitude[i] <- as.numeric(json_data$longitude)
}

df3 <- data.frame(Name, Latitude, Longitude)
```

Convert the data frame to a simple feature data frame and make an interactive map.
```{r}
library(sf)

projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
df.sf <- st_as_sf(df3, 
                  coords = c("Longitude", "Latitude"),
                  crs = projcrs)

library(tmap)

tmap_mode("view")

tm_shape(sfdf) +
  tm_dots() +
  tm_text(text = "Name")
```

Create a space-time object.
```{r}
library(stars)

d <- st_dimensions(location = df.sf$geometry, 
                   time = times)

df.st <- st_as_stars(list(Temperature = matrix(stack(df2)$values, ncol = length(times))), dimensions = d)
```






Choose an active WeatherSTEM site. Get raw data by clicking on <raw data> under current conditions.

One station: https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/apalacheepark/davis.json

Get the data into R.
```{r}
library(rjson)

loc <- "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/wtmoore/davis.json"
json_data <- fromJSON(file = loc)
```

Create a data frame
```{r}
ID <- json_data$davis_current_observation$DID
Name <- json_data$davis_current_observation$station_name
Latitude <- as.numeric(json_data$latitude)
Longitude <- as.numeric(json_data$longitude)
Day_Low_Time <- json_data$davis_current_observation$temp_day_low_time
Day_Low_Temp <- as.numeric(json_data$davis_current_observation$temp_day_low_f)

df <- data.frame(ID, Name, Latitude, Longitude, Day_Low_Temp, Day_Low_Time)
```

Create a station list.
```{r}
loc <- c("https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/apalacheepark/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/bucklake/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/canopyoaks/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/chaires/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/clc/davis.json",
#         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/chiles/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/conley/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/deerlake/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/desototrail/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/famudrs/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsus/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fortbraden/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsuesc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsursp/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/fsurez/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/gilchrist/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/hartsfield/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/hawksrise/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sullivan/davis.json", 
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/killearnlakes/davis.json",
          "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/leon/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/lincoln/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/maclay/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/netallahassee/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/oakridge/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/rickards/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/ruediger/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sail/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/sealey/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/success/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/swiftcreek/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/tcc/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/tallahasseenurseries/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/theedison/davis.json",
         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/woodville/davis.json"
#         "https://cdn.weatherstem.com/dashboard/data/dynamic/model/leon/wtmoore/davis.json"
)
```

Loop over the station list importing the data and making a data frame.
```{r}
ID <- rep(NA, length(loc))
Name <- rep(NA, length(loc))
Latitude <- rep(NA, length(loc))
Longitude <- rep(NA, length(loc))
Date <- rep(NA, length(loc))
Sunrise <- rep(NA, length(loc))
Day_Low_Time <- rep(NA, length(loc))
Day_Low_Temp <- rep(NA, length(loc))

for(i in 1:length(loc)){
  json_data <- fromJSON(file = loc[i])

    ID[i] <- json_data$davis_current_observation$DID
    Name[i] <- json_data$davis_current_observation$station_name
    Latitude[i] <- as.numeric(json_data$latitude)
    Longitude[i] <- as.numeric(json_data$longitude)
    Date[i] <- substring(json_data$observation_time_rfc822, 6, 16)
    Sunrise[i] <- json_data$davis_current_observation$sunrise
    Day_Low_Time[i] <- json_data$davis_current_observation$temp_day_low_time
    Day_Low_Temp[i] <- as.numeric(json_data$davis_current_observation$temp_day_low_f)
}
```

```{r}
( df <- data.frame(ID, Name, Latitude, Longitude, Date, Sunrise, Day_Low_Time, Day_Low_Temp) %>%
    mutate(Date = as.Date(Date, format = "%d %B %Y")) %>%
#    filter(Name != "DeSoto Trail Elementary WeatherSTEM") %>%
    filter(Date == as.Date("2020-12-12")) )
```

Convert the data frame to a simple feature data frame and make an interactive map.
```{r}
library(sf)

projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
sfdf <- st_as_sf(df, 
                 coords = c("Longitude", "Latitude"),
                 crs = projcrs)

library(tmap)

tmap_mode("view")

tm_shape(sfdf) +
  tm_dots() +
  tm_text(text = "Name")

tm_shape(sfdf) +
  tm_text(text = "Day_Low_Temp")
```

Get locations for interpolation.
```{r}
grid.sfc <- st_make_grid(sfdf,
                         n = c(20, 40),
                         what = "polygons")

tm_shape(grid.sfc) +
  tm_borders() +
tm_shape(sfdf) +
  tm_dots(col = "red")
```

Interpolate temperatures to a grid
```{r}
library(gstat)

interpolation <- idw(formula = Day_Low_Temp ~ 1,
                     locations = sfdf,
                     newdata = grid.sfc)

tm_shape(interpolation) +
  tm_fill(col = "var1.pred", alpha = .5)
```

Use projection first. Here we use Florida GDL Albers (EPSG:3087). This eliminates the warnings.
```{r}
sfdf <- st_transform(sfdf, 
                     crs = 3087)
grid.sfc <- st_make_grid(sfdf,
                         n = c(20, 40),
                         what = "polygons")
interpolation <- idw(formula = Day_Low_Temp ~ 1,
                     locations = sfdf,
                     newdata = grid.sfc)

tm_shape(interpolation) +
  tm_fill(col = "var1.pred", alpha = .5)
```

Interpolate to census tracts and correlation with median household income
```{r}
library(tidycensus)

median_hh_income <- get_acs(
  geography = "tract",
  variables = "B19013_001",
  state = "FL",
  county = "Leon",
  year = 2019,
  geometry = TRUE,
  show_call = TRUE
)

sfdf <- st_transform(sfdf, 
                     crs = 3087)
median_hh_income <- st_transform(median_hh_income, 
                                 crs = 3087)

interpolation <- idw(formula = Day_Low_Temp ~ 1,
                     locations = sfdf,
                     newdata = median_hh_income)

tm_shape(interpolation) +
  tm_fill(col = "var1.pred", alpha = .5)

cor.test(median_hh_income$estimate, interpolation$var1.pred)
```

Change in median household income
```{r}
median_hh_income <- get_acs(
  geography = "tract",
  variables = "B19013_001",
  state = "FL",
  county = "Leon",
  year = 2019,
  geometry = TRUE,
  show_call = TRUE
)
median_hh_income2015 <- get_acs(
  geography = "tract",
  variables = "B19013_001",
  state = "FL",
  county = "Leon",
  year = 2015,
  geometry = TRUE,
  show_call = TRUE
)

median_hh_income$estimate2 = median_hh_income2015$estimate

median_hh_income <- median_hh_income %>%
  mutate(PercentChange = (estimate - estimate2)/estimate2 * 100)



st_crs(median_hh_income) <- 4326
tmap_mode("view")

tm_shape(median_hh_income) +
  tm_fill(col = "PercentChange", alpha = .5)

```         