---
title: "Zach_Fires"
author: "Zach Law"
date: "3/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This file will be used to further analyze fire data bounded by the Apalachicola National Forest.


##Importing/creating boundaries of Apalachicola National Forest

Link to National Forest Boundary File: https://data.fs.usda.gov/geodata/edw/datasets.php?xmlKeyword=National+Forest+System+Land+Units

Link to zipfile containing a shapefile for all forest boundaries: https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip 

*Note: This link provides boundaries for every national forest boundary in the U.S.
Because we only want the boundary for the Apalachicola National Forest, the file was edited in esri arcpro and a shapefile containing only the Apalachicola National Forest boundaries was exported for use here.

```{r}
library(tidyverse)
library(sf)

anfbounds.sf <- read_sf(dsn = "forest boundaries",
                     layer = "ANFboundaries") %>%
  st_transform(crs = 3086)

st_crs(anfbounds.sf)
  
```
create a zoomable map
```{r}
library(tmap)

tmap_mode("view")

tm_shape(anfbounds.sf) +
  tm_borders()

```


Create window for apalacicola national forest boundary

```{r}
library(spatstat)
library(maptools)

anf.win <- anfbounds.sf %>%
  as_Spatial() %>%
  as.owin()

plot(anf.win)
```

Alternative method for importing forest shapefile using only R.

```{r}
library(tidyverse)
library(sf)

if(!"S_USA.NFSLandUnit" %in% list.files()){
  download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.NFSLandUnit.zip",
                "S_USA.NFSLandUnit.zip")
unzip("S_USA.NFSLandUnit.zip")
}

NF_Bounds.sf <- st_read(dsn = "s_USA.NFSLandUnit.shp") %>%
  st_transform(crs = 3086)
```

Filter shapefile for just Apalachicola National Forest. Table won't load and we don't know how to isolate the apalachicola forest boundary.


Window for all forest boundaries
```{r}
library(spatstat)
library(maptools)

nf.win <- NF_Bounds.sf %>%
  as_Spatial() %>%
  as.owin()

plot(nf.win)
```

##Importing Fire Data

```{r}
if(!"FL_Fires" %in% list.files()){
download.file("http://myweb.fsu.edu/jelsner/temp/data/FL_Fires.zip",
"FL_Fires.zip")
unzip("FL_Fires.zip")
}
FL_Fires.sf <- st_read(dsn = "FL_Fires") %>%
st_transform(crs = 3086)
```

Filter FL_Fires simple feature dataframe for only fires bounded by anf
```{r}
anf_fires.sf <- st_join(FL_Fires.sf, anfbounds.sf, join = st_within) %>%
  filter(NFSLANDU_2 == 'Apalachicola National Forest')
```


Create fire data into ppp object
```{r}
FL_fires.ppp <- FL_Fires.sf %>%
  as_Spatial() %>%
  as.ppp() %>%
  unmark()
```


Filter for fires only in the Apalachicola National Forest Boundary
```{r}
anf_fires.ppp <- FL_fires.ppp[anf.win]

anf_fires.ppp

anf_firesS.ppp <- rescale(anf_fires.ppp, s = 1000, unitname = "kilometer")

summary(anf_firesS.ppp)
plot(anf_firesS.ppp)
```
It is noted that a large amount of fires are clustered in the northeast boundary of the forest.

Note: We want to rescale the ppp object find the units.

```{r}
intensity(anf_firesS.ppp)

anf_firesS.ppp %>%
  density() %>%
  plot()
```



Look at fire data for the apalachicola national forest by month and year
```{r}
library(lubridate)
library(ggplot2)

anf_fires.sf %>% 
  ggplot(mapping = aes(x = month(DISCOVERY_))) +
    geom_bar() +
    facet_wrap(~ FIRE_Y)
``` 

View all apalachicola fires by month
Running this code we see that fires are most active during May and June and least active during December
```{r}
anf_fires.sf %>% 
  ggplot(mapping = aes(x = month(DISCOVERY_))) +
    geom_bar()
```

plot all may fires in the apalachicola boundary
```{r}
june_fires <- anf_fires.sf %>%
  filter(month(DISCOVERY_) == '6')
  
june_fires.ppp <- june_fires %>%
  as_Spatial() %>%
  as.ppp() %>%
  unmark()

june_fires.ppp <- june_fires.ppp[anf.win]

plot(june_fires.ppp)

```

plot apalachicola fires for the active fire period of June 2011

```{r}
june2011fire <- anf_fires.sf %>%
  filter(FIRE_Y == "2011") %>%
  filter(month(DISCOVERY_) == "6")

june2011fire.ppp <- june2011fire %>%
  as_Spatial() %>%
  as.ppp() %>%
  unmark()

june2011fire.ppp <- june2011fire.ppp[anf.win]

plot(june2011fire.ppp)

```



Plots based on fire size to note any trends
```{r}
anf_fires.sf %>% 
  ggplot(mapping = aes(x = FIRE_SIZE_)) +
    geom_bar() +
    facet_wrap(~ FIRE_Y)
```



All Florida fires visualized by year

Was 1996 a la nina year? Was 2010 an el nino year? 
```{r}
FL_Fires.sf %>% 
  ggplot(mapping = aes(x = month(DISCOVERY_))) +
    geom_bar() +
    facet_wrap(~ FIRE_Y)
```

##Plot Florida fires and drought for June 2011


Create window for Florida

```{r}
library(USAboundaries)

FL.sf <- us_states(states = "Florida") %>%
  st_transform(crs = st_crs(FL_Fires.sf))

FL.win <- FL.sf %>%
  as_Spatial() %>%
  as.owin()
```



All Florida Fires June 2011

https://droughtmonitor.unl.edu/Data/GISData.aspx

```{r}
june2011.sf <- FL_Fires.sf %>%
  filter(FIRE_Y == '2011') %>%
  filter(month(DISCOVERY_) == '6')

june2011.ppp <- june2011.sf %>%
  as_Spatial() %>%
  as.ppp() %>%
  unmark()

june2011.ppp <- june2011.ppp[FL.win]

plot(june2011.ppp)
```

Download drought shapefile for june
US Drought Monitor link to a list of shape files https://droughtmonitor.unl.edu/Data/GISData.aspx

Here the shape file for May 31, 2011 was used to represent drought for the period of June 2011

```{r}
if(!"USDM_20110531" %in% list.files()){
download.file("https://droughtmonitor.unl.edu/data/shapefiles_m/USDM_20110531_M.zip",
"USDM_20110531_M.zip")
unzip("USDM_20110531_M.zip")
}
drought062011.sf <- st_read(dsn = "USDM_20110531.shp") %>%
st_transform(crs = st_crs(FL_Fires.sf))
```


plot fires within the drought shapefile
why aren't colors showing up?

```{r}
drought062011.ppp <- drought062011.sf %>%
  as_Spatial() %>%
  as.owin()

drought062011.ppp <- drought062011.ppp[FL.win]

plot(june2011.ppp[drought062011.ppp])

```
```{r}
plot(drought062011.sf)
```





